### 2025-03-16 Dissertation/Paper/Front/01_coverpage.tex
Current Dissertation Progress

\begin{titlepage}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth, right]{Figures/logo.png}
\caption*{}
\label{fig:entropy} 
\end{figure}

\vspace{0.5in}

\centering
\Huge{\textbf{Dynamic Difficulty adjustment aided by Reinforcement Learning Agents in Real Time Strategy games. \\Also Second Line}}\\[2.0in]

\large{\textit{\textbf{Joshua James Polanszky}}} \\[0.2in]
\large{\textit{\textbf{Supervisor: Ms Maureen Cristina}}} \\[0.8in]

\vspace{1.5in}

\normalsize{\textbf{June - 2025}}\\[0.2in]

\normalsize{\textbf{A dissertation submitted to the Institute of Information and Communication Technology in partial fulfilment of
the requirements for the degree of BSc (Hons) Multimedia in Software Development}}\\[0.2in]

\end{titlepage}
\newpage

### 2025-03-16 Dissertation/Paper/Sections/01_Introduction.tex
Current Dissertation Progress

%=== CHAPTER ONE (1) ===
%=== INTRODUCTION ===

\chapter{Introduction}
\begin{spacing}{2.0}
%\setlength{\parskip}{0.2in}

In this section, \textbf{you}, the Student, are expected to state clearly:
%An alphabetcial list of items. Change the \alph* to \Alph* or \roman*
\begin{enumerate}[label=(\alph*)]
    \item the `problem’ or `question’ being researched;
    \item why this topic was chosen;
    \item what motivated the you to choose this topic;
    \item why did you investigate the topic the way you did;
    \item what problem did the you wish to explore;
    \item what is the context for the research?
\end{enumerate}
%Percentage amount
Percentage amount of words in section: 10 \% of Dissertation*

\section{Sub-chapter One}

Background goes here. Also you can put in some references~.

Another example of citations.

Here is a sample of table in \autoref{tabelsample}

\begin{table}[ht]
\centering
\arrayrulecolor[rgb]{0.596,0.596,0.596}
{\def\arraystretch{1.5}
    \begin{tabular}{|c|c|c|c|c|} 
    \hline

    \textbf{Age Groups}  & \textbf{Frequency} & \textbf{Percent} & \textbf{Valid Percent} & \textbf{Cumulative Percent}  \\[5pt]

    \hline 
    \textbf{16-20 years} & \textbf{100}       & \textbf{98.2}    & \textbf{98.2}          & \textbf{95.2}                \\ 
    \hline
    \textbf{21-25 years} & \textbf{5}         & \textbf{4.8}     & \textbf{4.8}           & \textbf{100.0}               \\ 
    \hline
    \textbf{Total}       & \textbf{105}       & \textbf{100.0}   & \textbf{100.0}         &                              \\
    \hline
    \end{tabular}
}
\arrayrulecolor{black}
\caption{Age of Participants}
\label{tabelsample}
\end{table}

Use  \texttt{$\backslash$newpage} to force start a new page.

\newpage

A very quick way to create tables in a point \& click environment is to use an online table generator\footnote{https://www.tablesgenerator.com}

\begin{figure}[ht]
\centering
\includegraphics[width=14cm, fbox]{Figures/tablesgenerator.jpg}
\caption{Create a Booktabs style table}
\label{fig:tablegenerator} 
\end{figure}

Use  \texttt{$\backslash$enquote} for double-quotes. \enquote{This is a sample quote.}

Also can try to refer to this image in \autoref{fig:boundingboxexample}. Notice that the \texttt{.eps} and \texttt{.pdf} format vector graphs are favoured, because:

\begin{enumerate}
    \item they can be zoomed-in to check the detail.
    \item text in such formats are search-able.
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=4in, fbox]{Figures/boundingbox.eps}
\caption{Bounding-box example of cars.}
\label{fig:boundingboxexample} 
\end{figure}

Try to insert a math equation as in \autoref{eq:euler}. If you wanna try the in-line mathematical, here is a sample $\alpha = \pi \cdot \frac{1}{\Theta}$.

\begin{equation}
\label{eq:euler}
    e^{ix}= \cos{x} + i \sin{x}
\end{equation}

%Also here is a sample for footnote and hyperlink url\footnote{\url{https://github.com/doem97}}.

When mention some file formats can use \texttt{music.mp3}, \texttt{latex.pdf}, etc.

\section{Sub-chapter Two}

Contrary to popular belief, Lorem Ipsum is not simply random text. It has roots in a piece of 
\todo{This is a todo note which appears in the margin}
classical Latin literature from 45 BC, making it over 2000 years old. Richard McClintock, a Latin professor at Hampden-Sydney College in Virginia, looked up one of the more obscure Latin words, consectetur, from a Lorem Ipsum passage, and going through the cites of the word in classical literature, discovered the undoubtable source. Lorem Ipsum comes from sections 1.10.32 and 1.10.33 of de Finibus Bonorum et Malorum (The Extremes of Good and Evil) by Cicero, written in 45 BC. This book is a treatise on the theory of ethics, very popular during the Renaissance. The first line of Lorem Ipsum, \enquote{Lorem ipsum dolor sit amet}.., comes from a line in section 1.10.32.

It is a long established fact that a reader will be distracted by the readable content of a page when looking at its layout. The point of using Lorem Ipsum is that it has a more-or-less normal distribution of letters, as opposed to using 'Content here, content here', making it look like readable English. Many desktop publishing packages and web page editors now use Lorem Ipsum as their default model text, and a search for 'lorem ipsum' will uncover many web sites still in their infancy. Various versions have evolved over the years, sometimes by accident, sometimes on purpose (injected humour and the like).

\section{Sub-chapter Three}
\todo[inline]{This todo note appears in line with the text. Todo notes are a great tool to leave comments or notes for self and can then be simply commented out.}
Vivamus eget odio tellus. Nam libero augue, eleifend molestie est eu, tempus vehicula magna. Sed dignissim imperdiet urna, non viverra risus blandit in. In lacinia aliquet leo, ut interdum ipsum ornare sed. Praesent ac fringilla justo. Vivamus pharetra non ipsum eget semper. In hac habitasse platea dictumst. Donec neque lectus, ultricies id massa posuere, sodales interdum ante. Sed et nunc vitae urna ornare porttitor nec vitae urna. Curabitur sit amet luctus urna. Nulla porta malesuada rutrum. Pellentesque vitae velit sed odio tincidunt iaculis. Nulla facilisi.

\newpage

\section{Sub-chapter Four}

To create a bulleted list you need to make use of the ''itemize" environment. A \LaTeX environment is a special section within a page where items as placed. An environment always has a ``\textbackslash{begin}" and a ``\textbackslash{end}". Items you can place in such an environment include:
\begin{itemize}
    \item enumerate -- used to create numbered/letter lists
    \item itemize -- creates a bullet list
    \item figures -- to add images or figures (.eps is the recommended format
    \item tables -- you can create them in tablesgenerator.com then copy the \LaTeX\\code and paste it.
    \item equations -- for those really neat looking mathematical formulae
    \item this list is not exhaustive...
\end{itemize}

\newpage
\section{Sub-chapter Five}
The following are some of the most common commands used during your writing. Some characters are reserved and cannot be directly written as in a normal word processor but need to be ``escaped". Then there are other useful commands such as adding a footnote, inserting a new line (to begin a new paragraph after a blank line, adding labels to items for cross-referencing, etc.

\begin{itemize}
    \item The following are special characters which require a \textbackslash{} before each character so the character is reproduced on the output. The tilde and exponent are even more special. 
    \begin{itemize}
        \item \textbackslash{\#}
        \item \textbackslash{\$}
        \item \textbackslash{\%} -- comment
        \item \textbackslash{\&}
        \item \textbackslash{\_}
        \item \textasciitilde -- type \textbackslash{textasciitilde}
        \item \textasciicircum -- type \textbackslash{textasciicircum}
    \end{itemize}
    \item \textbackslash{textbf} -- bold font
    \item \textbackslash{textit} -- italics
    \item \textbackslash{underline\{words to underline\}}
    \item \textbackslash{emph} -- emphasized text. If used within an italicized text, the output is normal font. If used by itself, text in its argument is italicized.
\end{itemize}
An example highlighting the use of \textbackslash{emph}.
\begin{itemize}
    \item An example of emphasized text with \emph{these words emphasized} within normal font.
    \item \textit{This is an example where \emph{these three words} are emphasized within italicized font}
\end{itemize}

\end{spacing}
%=== END OF CHAPTER ONE ===
\newpage

### 2025-03-16 Dissertation/Paper/Sections/02_LiteratureReview.tex
Current Dissertation Progress

%=== CHAPTER TWO (2) ===
%=== Literature Review ===

% TODO: Add the .bib references so that they are listed properly.

\chapter{Literature Review}
\begin{spacing}{2.0}
	%\setlength{\parskip}{0.2in}

	\section{Chapter Introduction}

	Video games are constantly evolving to provide more engaging and immersive experiences to players, be it by improved graphics fidelity, better story telling, or by increasing the challenge.
	When it comes to the balancing difficulty to player skill, traditional difficulty techniques often fail to properly adapt to the vast range of player skill levels, and the difference on how fast certain players learn.
	Dynamic Difficulty Adjustment (DDA), along with Reinforcement Learning (RL) agents emerge as solutions to adapt game difficulty in real time based on player performance, along with decreasing predictability and
	increasing the challenge players face, enhancing player immersion, motivation and increasing replayability \cite{grech_creating_2023} \cite{mercieca_evaluating_2023} \cite{attard_analysing_2021}. This section explores and discusses previous works and their results regarding DDA and RL agents in video games
	to identify the current state of the art and to provide a foundation for the research conducted in this thesis.

	% make dda come after Unity-ML.
	\section{Dynamic Difficulty Adjustment (DDA)}

	Dynamic Difficulty Adjustment (DDA) refers to a game design approach where a set of systems dynamically alter the difficulty of a game through modifying game parameters such as AI behaviour,
	player and enemy stats, or environmental factors in real time to match player's skill. These systems target maintaining player engagement to be in line with Csikszentmihalyi's Flow Model, also known as the
	"flow state", a mental condition that is characterised by its ability to keep players deeply immersed and focus on gameplay, avoiding boredom or frustration \cite{grech_creating_2023} \cite{mifsud_utilizing_2021} \cite{laus_dynamic_2022} \cite{vang_impact_2022}.

	% Mention the flow state a bit more
	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/Flow_State.png}
		\caption{Flow State. \cite{sepulveda_exploring_2019}}
		\label{fig:flowstate}
	\end{figure}

	One of the fundamental aspects of DDA is its adaptability to player performance, which allows it to better accomodate the diverse range of player skill levels and learning speeds, avoiding the frustration that
	can occur with traditional fixed difficulty settings. Research has shown that several DDA implementations, such as ramping (rDDA) and player-controlled adjustments (pDDA), can significantly improve player experience.
	And recently, these approaches on DDA have been combined with AI techniques and algorithms to create more sophisticated and responsive systems that can adapt to player performance in real time, such as
	Deep Reinforcement Learning (DRL), Fuzzy Logic, Decision trees and Monte Carlo Tree Search (MCTS) \cite{attard_analysing_2021} \cite{chrysafiadi_fuzzy_based_2023}.
	% Mention the specific data needed for DDA to work in these types of games \cite{grech_creating_2023}

	\subsection{Determining Player Skill}

	Determining player skill is a crucial aspect of DDA, as if the DDA system is innaccurate, it can lead to the opposite effect, where the player will become more bored or frustrated, and can even lead to player's
	leaving the game. To develop effective DDA systems, it is crucial that developers take care to choose the variables that best determine the player's skill level. In the work done by Hunicke {31}, the Hamlet DDA system was used to
	closely monitor the core inventory, including the players health, shielding, ammunition, and weapons, adjusting the difficulty of the level based on the player's items. In a combat encounter, the DDA system was
	developed to see rapid depletion of items as a sign of player struggle, and therefore the system would intervene by modifying the level to aid the player {31}.

	\section{Static Game Difficulties}

	Currently, most games use static difficulty settings, where players choose a static difficulty level before starting the game.
	There are many reasons for this, such as the simplicity of implementation, the ability to control the player experience
	and progression, and that in the past players did not like the instability of dynamic difficulty systems, which with older algorithms
	did not have access nor the ability to process the complex data needed for DDA to work in these types of games \cite{attard_analysing_2021}

	\subsection{Fixed Difficulties}

	These fixed difficulties rely on pre-set levels (e.g., easy, medium, hard) chosen by players before playing, and suffer from
	being rigid and unable to adapt to player performance, skill levels or evolving gameplay dynamics \cite{attard_analysing_2021} \cite{laus_dynamic_2022}.
	% Discuss more

	\subsection{Difficulty Selection}

	With difficulty selection, players can change the difficulty during gameplay, helping in reducing the frustration or boredom that
	comes from fixed difficulties, and helps in keeping players engaged for longer periods \cite{attard_analysing_2021} \cite{laus_dynamic_2022}. However, this approach still requires players to manually adjust the difficulty,
	which can be cumbersome and disrupt the flow of the game, and are still rigid in nature, leading to the difficulty always being either too easy or too hard \cite{attard_analysing_2021} \cite{laus_dynamic_2022}.
	% Discuss more

	\section{Dynamic Difficulty Adjustment Compared with Static Difficulties}

	As discussed, dynamic systems outperform static ones when it comes to keeping to Csikszentmihalyi's Flow Model, helping reduce player boredom and frustration while enhancing both player immersion and replayability,
	thus keeping players engaged for longer periods \cite{grech_creating_2023} \cite{attard_analysing_2021} \cite{mifsud_utilizing_2021}. The main drawbacks of DDA come from its more complex development process, the need for more data and computational resources, and the potential for it to
	become more boring and frustrating than fixed difficulties if not implemented correctly, and given an adequate amount of real-time data \cite{attard_analysing_2021} \cite{laus_dynamic_2022}.

	\subsection{Game Immersion and replayability}

	Studies such as those done by \cite{mercieca_evaluating_2023} \cite{attard_analysing_2021} \cite{mifsud_utilizing_2021} \cite{laus_dynamic_2022} \cite{vang_impact_2022} confirm that DDA significantly improves player immersion by keeping players in a balanced challenge zone, keeping them both motivated and better entertained, resulting in
	a better player experience. Furthermore, the dynamic nature of these systems motivate players to revisit games, as the change in difficulty and challenges they present the player scale with the player's skill,
	keeping the game feeling fresh and engaging where fixed difficulties could not \cite{grech_creating_2023} \cite{attard_analysing_2021} \cite{mifsud_utilizing_2021} \cite{laus_dynamic_2022} \cite{sepulveda_exploring_2019}.

	\section{AI with DDA Algorithms}

	Studies have also shown that in recent time, AI has played a pivotal role in improving on the implementation of DDA systems by allowing better decision making and data processing based on player performance metrics.
	Such models are Reinforcement Learning (RL), as shown in the work done by \cite{mercieca_evaluating_2023} \cite{mifsud_utilizing_2021} \cite{zheng_dynamic_2024}, fuzzy logic \cite{chrysafiadi_fuzzy_based_2023}, Monte Carlo Tree Search (MCTS) \cite{demediuk_monte_2017} \cite{soemers_enhancements_2016}, and Decision Trees \cite{sejrsgaard_jacobsen_dynamic_2011}, also known as Behaviour trees.
	These different AI algorithms have helped enable more nuanced and responsive DDA implementations, offering players what feels like a more tailored experience \cite{attard_analysing_2021} \cite{vodopivec_monte_2017} \cite{chrysafiadi_fuzzy_based_2023}.

	\section{Adaptive Gameplay AI Algorithms}

	\subsection{Deep Reinforcement Learning (DRL) Algorithm}

	% TODO: Go more in depths about PPO, since it is the most common algorithm used in games.
	RL algorithms, and in turn Deep Reinforcement Learning (DRL) algorithms use neural networks to enable AI agents to learn and adapt to the situation presented to them. This is done through multiple methods, with the
	most common being Proximal Policy Optimisation (PPO), the State Action Reward State Action (SARSA) and Q-learning algorithms, which have been used to great effect in games like \cite{bin_ramlan_implementation_2021} \cite{metz_evaluation_2020}. They work by placing the agent in a controlled environment,
	where they learn through trial and error, being rewarded for good decisions and penalised for poor ones. This causes the agent to learn the optimal strategy for the situation presented to them,
	and can be used not only as opponents or allies in games, but also as a way to dynamically adjust game difficulty in real time, particularly in fighting games and other high-interaction environments,
	as seen in the works done by \cite{mercieca_evaluating_2023} \cite{mifsud_utilizing_2021} \cite{zheng_dynamic_2024}.
	% Discuss these works a lot more in the RL Agents section, which will be moved above this.

	% Mention the flow state a bit more
	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/SarSa.png}
		\caption{Agent training environment interaction in RL. \cite{grech_creating_2023}}
		\label{fig:rl_training}
	\end{figure}

	% Should likely expand, but only once the latter part about RL-Agents is done. On top of that, too much time has been spent fine tuning this, and as such it is likely better to move on to the other sections and then come back to it.
	% Q-Learning and Proximal Policy Optimisation (PPO) algorithms, which have been used to
	% Algorithms such as Proximal Policy Optimization (PPO) have been effective in tailoring game difficulty in real time, particularly in fighting games and other high-interaction environments \cite{bin_ramlan_implementation_2021} \cite{metz_evaluation_2020}.

	\subsection{Fuzzy Logic}

	In the study done by Chrysafiadi, et al. \cite{chrysafiadi_fuzzy_based_2023}, Fuzzy logic systems were employed to dynamically adjust game difficulty to match the player's skill level, with the goal of maintaining player engagement while teaching them HTML programming.
	Fuzzy logic systems are particularly effective in such real-time adjustments, as they require less training data when compared to ML models. Moreover, trhese systems offer immediate real-time adjustments to better align with player performance.
	The results of this study show that player dropouts decreased significantly, and player engagement and motivation increased, and even performed better than other groups, proving that fuzzy-based DDA can enhance
	educational outcomes and help keep players in the Flow State with reduced data \cite{chrysafiadi_fuzzy_based_2023}.
	% Perhaps add statistics and more information on the study?

	\subsection{Monte Carlo Tree Search (MCTS)}

	MCTS works by exploring different game states in the form of simulations to optimise decision-making, and is done by building a search tree iteratively, balancing exploration and exploitation to find the best
	move \cite{vodopivec_monte_2017} \cite{demediuk_monte_2017}. In the work done by Demediuk, et al. \cite{demediuk_monte_2017}, MCTS was used to enhance player engagement by modifying AI strategies to maintain balanced matches, with variants like Reactive Outcome-Sensitive Action Selection (ROSAS).
	ROSAS is reactive in nature, which can make it feel unnatural, whilst Proactive OSAS (POSAS), which is more proactive and can feel more natural, at the cost of precision. In the work done by Soemers, et al. \cite{soemers_enhancements_2016},
	different MCTS optimisations were made, such as Progressive History (PH) and N-Gram Selection Technique (NST), where PH biases the selection step towards actions that have been successful in the past, and NST
	biases the play-out step towards actions that have been successful in the past, with the results showing that when combined these greatly improve the algorithm's performance. Another improvement was Deteministic
	Game Detection (DGD), which modifies MCTS behaviour to optimise for deterministic games by using alternative evaluation strategies, which was effective when combined with other enhancements \cite{soemers_enhancements_2016}.
	What these studies show, is that when it comes to more complex and deterministic games, MCTS can be a powerful tool to enhance player engagement both as an opponent, as well as a tool to dynamically adjust
	game difficulty.

	\subsection{Decision \& Behaviour Trees}

	% TODO: Split Decision and behaviour trees based on extra research done.
	Decision trees are popular tools in data analysis, and are used for classification and prediction due to their intuitive structure, which include nodes (root, internal, leaf) and branches \cite{song_decision_2015} \cite{mienye_survey_2024}.
	The framework around how decision trees work when it comes to games varies, as not only can it reduce code repetition in game development, both for complex and simple games \cite{vuong_artificial_2020}, but its structure allows for easy
	interpretation and visualisation of the algorithm's decision making process, making them particularly useful for adapting game features dynamically, as seen in the work done by \cite{sejrsgaard_jacobsen_dynamic_2011}.
	Their approach involved examining two distinct approaches, one using predefined behaviour trees of varying difficulties, and another employing a single tree with dynamic adjustment of probabilities to alter difficulty.
	The first approach struggled when it came to flexibility, as it could not adapt to player performance, which was in stark contrast with the second approach, which was able to optimise behaviour probabilities
	and display superior adaptability, helping to keep the player in the Flow State, and maintain player engagement \cite{sejrsgaard_jacobsen_dynamic_2011}. Along with this, decision trees have been shown to be even more effective when combined with
	other algorithms, such as RL, which help greatly in improving their efficiency and scalability in complex gaming environments, or even with state machines in more simple environments \cite{mcquillan_survey_nodate} \cite{sejrsgaard_jacobsen_dynamic_2011}.

	% Here

	\section{Procedural Content Generation}

	Procedural Content Generation (PCG) is a technique used in game development to create game content algorithmically, leveraging algorithms to dynamically create and add content such as levels, environments,
	and items to games that both help in reducing development time and costs, but also increase replayability and can easily be combined with DDA systems to create more engaging and immersive experiences \cite{laus_dynamic_2022}.
	When paired not only with DDA, but also with AI agents, such as those created by \cite{grech_creating_2023}, PCG was used to constantly modify the best possible action that the player and the agent could take, which in turn
	not only forced the player to adapt to the changing environent, but also helped keep the player feeling challenged and engaged, avoiding frustration since the opponent was also having to adapt in a human like way,
	helping in keeping the player in the Flow State.
	% This needs to be cut down and made better

	% Should finite state machines be included here?

	\section{Unity ML-Agents and how they work}

	Unity ML Agents is a toolkit that enables developers to more easily create and integrate intelligent agents within Unity environments using Machine Learning (ML) techniques.
	The package allows users to transform any scene into a training enviroment where agents can be trained using algorithms such as RL, Imitation Learning and neuroevolution, by allowing agents to observe their surroundings, take actions and receive rewards,
	as well as specifying the behaviours that these agents can take \cite{raut_unity_2024} \cite{unity_ml_agents}.

	\subsection{Finite State Machines}

	Traditionally, and even to this day the most common way to implement AI in games is through Finite State Machines (FSM), which are used to describe a relationship between a set of states, where transitions
	between these states are determined by condition, with actions being performed based on the current state \cite{grech_creating_2023}. FSMs are useful in simple environments, however as the number of actions and states are limited, However
	as the number of states increase, so does the complexity of the FSM, making it difficult to manage, debug and maintain \cite{grech_creating_2023}.

	\subsection{RL Agents}

	As mentioned in an earlier section, RL agents are AI agents that learn through trial and error, being rewarded for good decisions and penalised for poor ones. RL-Agents trained using Unity ML-Agents have more
	access to the environment and its data, and this has been used in the work done by Grech \cite{grech_creating_2023}, where he trained an RL-Agent to act as an opponent in a Real Time Strategy (RTS) game, taking multiple stages of the agent
	in training for the different difficulties. There were also different RL-Agent models such as the Proximal Policy Optimisation (PPO) and others that were used to act as enemies in fighting games \cite{bin_ramlan_implementation_2021}, bosses \cite{metz_evaluation_2020},
	and even for racing games \cite{berta_development_2024}. In these works, along with those done by \cite{borg_investigating_2020} \cite{zhasulanov_enhancing_2024} \cite{raut_unity_2024}, it was revealed that RL-Agents have shown many times to be an effective way to make a more unpredictable, life-like and challenging
	opponent for player, significantly boosting their experience and engagement.

	\subsection{RL Agents compared with Finite State Machines}

	The main differences between FSMs and RL-Agents are their development time and costs, as FSM's are easier and cheaper to implement, but are then limited in their complexity and scalability, where as RL-Agents
	take more time and resources to develop, but solve the issues with FSMs being their complexity, scalability, and the simple fact that when a player fights an FSM enough, they will learn the patterns and states,
	making the game predictable and boring \cite{grech_creating_2023} \cite{bin_ramlan_implementation_2021} \cite{metz_evaluation_2020}. This life-like aspect to RL-Agents can, and have been shown to greatly improve game replayability and immersion, keeping the player in the flow state for longer
	and providing a superior experience to the player \cite{grech_creating_2023} \cite{bin_ramlan_implementation_2021} \cite{metz_evaluation_2020}. With all of this however, it is important to not that if an RL-Agent is not implemented or trained correctly, it can lead to the opposite effect,
	where an opponent is too difficult, leading to player frustration, or too easy, leading to player boredom \cite{grech_creating_2023} \cite{bin_ramlan_implementation_2021} \cite{metz_evaluation_2020}.

	\subsection{Practical Applications}

	As mentioned in the previous sections, RL-Agents have been used in a variety of games, from RTS games to fighting games and now even racing games. The practical applications of RL-Agents in games are vast,
	and as hardware capabilities improve, and our understanding and training techniques improve, the potential for RL-Agents in games is almost limitless. The ability to create more engaging, immersive and challenging
	opponents for players, as well as allies, and the ability to easily integrate them with DDA systems, PCG and other game development techniques to make games more replayable and engaging is a powerful tool for developers
	to create better games \cite{grech_creating_2023} \cite{borg_investigating_2020} \cite{zhasulanov_enhancing_2024} \cite{raut_unity_2024} \cite{metz_evaluation_2020}

	\section{Chapter Overview}

	This chapter has provided an overview on the current state of the art in Dynamic Difficulty Adjustment (DDA) and Reinforcement Learning (RL) agents in video games, and how individually they have been used to
	improve upon more static techniques, such as fixed difficulties and Finite State Machines (FSMs). It has shown that integrating AI techniques has helped in improving the accuracy and adaptability of DDA systems,
	and the latest advancements in RL-Agents, including the Unity ML-Agents package, have shown just how revolutionary these techniques are at improving player engagement and immersion, as well as game replayability
	when these techniques are implemented.
	% Should more sources be referenced since this is more of a summary of the chapter?

	% TODO: Some parts have not been added to this from the word document, and some information is currently missing, however too much time has been spent on this draft so it is best to move on to the next section
	% and wait for feedback and then come back to this section to make the necessary changes.
	% On top of that, due to there being too much information not being condensed properly, the analysis is not as deep as it should be, and this should be improved in the next draft following recommendations.

	%=== END OF CHAPTER TWO ===
\end{spacing}
\newpage

### 2025-03-16 Dissertation/Paper/Sections/03_Methodology.tex
Current Dissertation Progress

%=== CHAPTER THREE (3) ===
%=== (Actual work done and contribution, including literature survey) ===

\chapter{Research Methodology}
\begin{spacing}{2.0}
	%\setlength{\parskip}{0.2in}
	%  (Actual work done and contribution, including literature survey)

	\section{Chapter Introduction}

	% Mention the aim of the study and reference the research questions, using the SOI.

	% Create mini summary of selected research papers and notes to better align the research questions and methodology with the literature review.

	% \section{Nature of Research}
	% % Should I discuss this in its own section, or just include my own notes and split it in the introduction and especiialy discuss it % in the data collection,
	% % as it is more relevant to the data collection and evaluation sections and is covered more in depth there.
	%    \subsection{}

	% Should this just be called the Research Pipeline?
	\section{Conceptual Framework}

	Figure \ref{fig:pipeline} outlines the methodology pipeline for this study's research, with it being divided into 4 major sections.
	The first section consisted of decided on the prerequisites and tools required for the development of the prototype, along with developing the
	traditional game mechanics and State Machine opponent. The second section consisted of setting up the training scene and Unity mL-Agents plugin,
	training, and evaluating the RL-Agent. The third section involved developing the data collection and developing the DDA system using behaviour-trees,
	which were decided over the other options due to their ease-of-use, and how well they are at representing the AI's decision making process.
	Finally, the prototype was evaluated by the participants, with the data collected being analysed and compared to previous research to answer the research questions.

	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/ResearchPipeline.png}
		\caption{Diagram representing the methodology pipeline.}
		\label{fig:pipeline}
	\end{figure}

	\section{Prerequisites and Tools}

	Before starting development on the prototype, a github repository for the project was created, and Jetbrains Rider was used as the IDE for the project.
	The prototype was developed using Unity 6000.0.31f1, and utilised several Unity packages, assets and plugins, which are listed below:

	% Do I list them here or reference an appendix?

	Along with the built-in Unity packages, the project also used the Unity ML-Agents package which was cloned off the Unity ML-Agents github repository, and the Unity ML-Agents Toolkit, which was cloned off the Unity ML-Agents Toolkit github repository,
	and then imported into the project as a Unity package along with the ML-Agents Extensions packages. The version of ML-Agents used in the prototype
	was version 22, the latest version at the time of development.

	\section{Prototype Outline \& Implementation}

	% Likely will need to split this into an outline, with the State Machine being a subsection, and the RL-Agent and DDA system being their own sections.

	A Prototype game was developed to implement and test the RL-Agent and DDA system for their effectiveness in creating a dynamic and adaptive game experience.
	The core gameplay will consists of traversing the game world in a third person perspective, with the goal to mine and deposit as many ores as possible.
	There are 3 main mechanics which for the core gameplay loop of the game, the scoring system, the mining system, and the inventory system. The mining system is responsible for the player's ability to mine ores,
	the scoring system is responsible for tracking the player's score and awarding them points based on the types or ores deposited, and the inventory system is responsible for forcing the player to carefully
	think about their inventory space and which ores to mine and when to go deposit. The ores and terrain are procedurally generated, with some small obstacles which can force the player to take a slightly longer path.

	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/Gameplay.png}
		\caption{Screenshot of gameplay showing the gathering of resouces.}
		\label{fig:gameplay}
	\end{figure}

	The main challenge the player will face is the AI opponent which has the same goal and capabilities of the player. The objective for both of them is to gather as many points as possible before the game ends.
	If the player has more points at the end, they win, if its the same, it is a draw, and if the AI has more points, the player loses. The AI opponent will be different for each stage of the test, with the first
	being a State Machine, and the second being an RL-Agent. More information regarding the test stages and how data will be gathered and analysed will be discussed in a later section.

	\subsection{Ores}

	There are three types of ores in the game, Copper, Silver, and Gold. Each ore is visually distinct, as can be seen in Figure \ref{ores} and have a different chance of spawning, as listed in Table \ref{ore_chance_table}.

	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/Ores.png}
		\caption{Screenshot showing the different ores.}
		\label{fig:ores}
	\end{figure}

	\begin{table}[ht]
		\centering
		\renewcommand{\arraystretch}{1.2} % Adjust row spacing
		\setlength{\tabcolsep}{8pt} % Adjust column spacing
		\begin{tabular}{|c|c|c|c|}
			\hline
			\textbf{Difficulty} & \textbf{Copper Spawn \%} & \textbf{Silver Spawn \%} & \textbf{Gold Spawn \%} \\
			\hline
			Easy                & 60\%                     & 30\%                     & 10\%                   \\
			\hline
			Medium              & 70\%                     & 25\%                     & 5\%                    \\
			\hline
			Hard                & 75\%                     & 20\%                     & 5\%                    \\
			\hline
		\end{tabular}
		\caption{Chance of spawning for each ore.}
		\label{ore_chance_table}
	\end{table}

	They also take different amounts of time to mine, inventory space, and points awarded when deposited, as listed in Table \ref{ore_table}.

	\begin{table}[ht]
		\centering
		\renewcommand{\arraystretch}{1.2} % Adjust row spacing
		\setlength{\tabcolsep}{8pt} % Adjust column spacing
		\begin{tabular}{|c|c|c|c|}
			\hline
			\textbf{Resource Type} & \textbf{Score} & \textbf{Mining Time (swings)} & \textbf{Inventory Space Taken} \\
			\hline
			Copper                 & 2              & 3                             & 1                              \\
			\hline
			Silver                 & 5              & 5                             & 3                              \\
			\hline
			Gold                   & 10             & 8                             & 5                              \\
			\hline
		\end{tabular}
		\caption{Ore statistics.}
		\label{ore_table}
	\end{table}

	\subsection{Traditional State Machine}

	The State Machince will be a simple AI opponent using the Unity navmesh system to navigate the game world. The AI will have 3 states, a mining state, a depositing state, and a travelling state.
	After completing a task, the AI will use a Physics OverlapSphere to search for the best ore within the vicinity, and once the best option is found, the AI will navigate to the ore and mine it.
	The speed at which the AI travels, and the radius it searches for ores will be adjusted based on the difficulty selected, allowing for a faster and more effective opponent on higher difficulties.
	Once the inventory is full, the AI will navigate to the closest deposit point to deposit the ore and gain points.

	\subsection{RL Agent}

	The Reinforcment Learning Agent will be trained using the Unity ML-Agents plugin, and will use Proximal Policy Optimisation (PPO) as the training algorithm, since it is the most commonly used algorithm due to
	its effectiveness at training agents in complex and continuously changing environments, the main RL training algorithm for the Unity ML-Agents framework, and the one used in the works done by \cite{grech_creating_2023},
    \cite{bin_ramlan_implementation_2021}, \cite{berta_development_2024}, and \cite{raut_unity_2024}.
	The agent will then be placed instead of the State-Machine as the opponent the player will face. More information on the training and evaluation will be discussed in a later section.

	\subsection{DDA System}

	This section will be created over time while the prototype and DDA system are being developed and tested.

	\section{RL-Agent opponent}

	\subsection{Agent Observations}

	To understand the game enviroment, the agent is supplied with several variables to observe via the Unity ML-Agents plugin. These variables are all normalised numerical values between 0 and 1, and are as follows:

	\begin{itemize}
		\item The type of ore currently being obesrved.
		\item How much of the inventory the ore will take up.
		\item The agents normalised position from the ore.
		\item The player's normalised position from the ore.
		\item The deposit point's normalised position from the ore.
		\item The normalised amount of inventory space left.
		\item The normalised amount of time left.
	\end{itemize}

	\begin{figure}[ht]
		\centering
		\includegraphics[width=5in, fbox]{Figures/Observations.png}
		\caption{Screenshot showing the Agent's observations.}
		\label{fig:observations}
	\end{figure}

	\subsection{Agent Decisions \& Actions}

	The Agent has 3 decisions it can make, and 2 actions it can take. The decisions are between travelling to and mining the ore being observed, travelling to the deposit point, or observing the next ore.
	The travel to \& mining action involves setting the ore's position as the NavmeshAgent's destination, waiting until the agent arrives, and then mine the ore. Sometimes due to obstacles and the agent's pathfinding,
	the agent may get stuck, which the program will detect very quickly by requesting the agent to decide on another ore. While this is not ideal, it is an uncommon bug with the NavmeshAgent.
	The other action the agent can take is travelling to the deposit point, which acts similarly to the travel to \& mining action, but instead of mining, the agent will just observe and make another decision once it arrives.

	% NOTE: Add flowchart of the agent's decision making process here.

	% TODO: Expand on the agent's observation and decision making.

	\subsection{Agent Rewards \& Penalties}

	A simple reward and penalty system was used to train the agent, since the enviroment is simple and the agent's goal is clear. The focus was on rewarding the agent for mining and depositing ores,
	and penalising them for travelling for too long.

	% TODO: Show the final rewards and penalty systems that will be used

	\subsection{Agent Training}

	\subsection{Model Selection \& Evaluation}

	\section{Data Collection}

	% Should the participants come before the type of data?
	% Should i mention the tests and how the data will be gathered after the sub sections, and use the sub sections to explain the type of data, why it is being collected and who it is being collected from, or should this come before?

	It was decided to use a mixed-methods approach for the evaluation of the prototype, with the majority of the data being quantitative.
	This was chosen to better match \cite{grech_creating_2023} \& \cite{mercieca_evaluating_2023}'s approaches, allowing for an in-depth comparison and analysis of results,
	while allowing for a more detailed understanding of the player's experience and recommendations. Quantitative data will be collected through Unity Analytics, allowing data to be extracted from gameplay sessions,
	as well as close ended questionnaires. Quantitative data will be collected through open ended questions in the questionnaire.

	\subsection{Quantitative Data}

	To better understand the performance of the agent and DDA system against a human player, and to compare the two against State Machines, important statistics throughout the game will be tracked, and saved to a database.
	This data will include the player and enemy action logs, the game summary, and the performance summary.

	\subsubsection{Player and Enemy Action Logs}

	These logs, as seen in \ref{fig:logs} will show all note-worthy actions taken by the player and the enemy, detailing the time, action and any relelvant information, such as the time left and score, and will be shown in chronological order.
	This data will be used to analyse the player's and enemy's behaviour and decision making throughout the game.

	\subsubsection{Game \& Perfomance Summary}

	The game summary will show some of the less note-worthy actions taken by the player and enemy, such as time spent mining and travelling. It will also show the final score and the difficulty selected.
	Due to the DDA system in place, some of the statistics will include logs of difficulty adjustments, and the final difficulty the DDA system selected. The performance summary will be used to capture the
	game's performance by tracking metrics such as frame-rate, CPU Usage and memory usage.

	\subsection{Qualitative Data}

	To better understand the player's experience, as well as some pain points of the game, the questionnaires will have some open ended questions, asking the player to better describe their experience,
	and to provide feedback and any recommendations that they might have. This data will help expand upon the metrics gathered from the quantitative data, and will be used to better answer the research questions.

	\subsection{Participants}

	To really gauge the effectiveness of RL-Agents and DDA systems in creating a dynamic and adaptive experience, a large diverse group of participants will be required.
	% TODO: Add more info about the number of participants and such instead of saying large and diverse.
	By having players of different ages and experience, the data collected will be more varied and allow for a more in-depth analysis of the data. This is inline with the research being conducted, as the goal is to
	identify places where AI is worth implementing, such as more difficult competitive games, or games with a more casual audience.

	\subsection{Tests}
	% Mention that they will agree to the notice found at the end of the SOI
	The tests will be conducted in two stages. First, a close-ended questionnaire will be taken by the participants to gather basic information anonymously, such as age and gaming experience. Once complete,
	the participants will be split into two groups, ensuring that both groups have a similar distribution of age and experience, and they will both play the traditional segment of the prototype,
	with traditional difficulty settings and a State Machine as the opponent. In the second stage, the first group will then play against the RL-Agent, whereas the second group will play against the RL-Agent
	with the DDA system in place. After each playtest, the participants will be given a questionnaire to describe their experience and provide feedback, with both closed and open-ended questions. As mentioned earlier,
	data from the playtests will also be gathered through Unity Analytics.

	\section{Evaluation}

	The data collected from the playtests and questionnaires will be analysed and cross-referenced to find any patterns or correlations between the data.
	The quantitative parts of the data will be analysed using T-Test analysis, and the qualitative data will be analysed using thematic analysis. The results will then be compared to previous research,
	and then be used to answer the research questions, and to prove or disprove the hypothesis.

	\section{Chapter Overview}

	%=== END OF CHAPTER THREE ===
\end{spacing}
\newpage

### 2025-03-16 Paper/main.tex
Added 2 col ieee template for RD2 assignment

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Title Goes Here}

\author{\IEEEauthorblockN{Joshua Polanszky}
\IEEEauthorblockA{\textit{Institute of Information Communication Technology} \\
\textit{Malta College of Arts Science and Technology}}
Paola, Malta
}

\date{Date Goes Here}

\maketitle

\begin{abstract}
abstract
\end{abstract}

\begin{IEEEkeywords}
Keywords
\end{IEEEkeywords}

\section{Introduction}

Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}

\section{Literature Review}

Literature review goes here

\section{Research Methodology}

Research Methodology goes here

\section{Findings}

Findings go here

\section{Conclusion}

Conclusion goes here:

\bibliographystyle{ieeetr}
\bibliography{references}
\end{document}

### 2025-03-16 Paper/main.tex
Section 1 of Task 1 almost done. Using Zotero for literature management, same as dissertation.

@@
 
-\title{Title Goes Here}
+\title{Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience}
 
@@
 
-\date{Date Goes Here}
+\date{\today}
 
@@
 
-Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}
+% Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}
+% TODO: Restructure a bit
+\subsection{Theme and Topic Rationale}
+% TODO: Cite the papers taken from the dissertation that are relevant to the topic, since the topic is similar to the disseration and as such, I already have the info i need without re reading.
+
+The Theme chosen is Decision-Making AI for Real-Time Strategy (RTS) Games, and will focus on comparing Finite State-Machine (FSMs) AI opponents traditionally used in games, against Machine Learning (ML) opponents, specifically
+Reinforcement Learning (RL), and their impact on player experience.
+
+Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
+While extensive research has been conducted on the topic, few have compared the experience these AI opponents provide to players, and focused more on the pure performance of the RL agent, rather than the
+player experience.
+This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational
+and development cost of implementing RL agents is justified by the improvement in player experience.
+
+\subsection{Positioning and Research Onion}
+
+This research will be positioned within the gap of player experience found in literature, and will help in provided a better understanding on the role RL agents will play in the future of RTS games, building off of the
+work done by \cite{vinyals_grandmaster_2019} in their work on AlphaStar.
+
+As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm, following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.45\textwidth]{../Images/Research_Onion.PNG}
+    \caption{Research Onion}
+    \label{fig:research_onion}
+\end{figure}
+
+\subsection{Background to the Research Theme}
+% TODO: Add references to papers from dissertation, and maybe expand
+Game AI has evolved significantly over the years, especially in RTS games.
+Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making.
+These FSM-based approached
+are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patters of the AI.
+
+More recently, RL has emerged as an alternative AI approach, taking advantage of advancements in ML and computer hardware.
+In games such as AlphaStar \cite{vinyals_grandmaster_2019}, RL agents were able to
+demonstrate adaptive and human-like behaviour, providing a more challenging and engaging experience for players.
+Despite all of this, the implementation of RL in commercial games remains limited due to
+the high computational cost, long training and development times, and added complexity.
+
+This further proves the need for research in this area, and in evaluating if the benefits of RL agents in RTS games are worth the cost compared to traditional FSMs.
+
+\subsection{Hypothesis}
+
+Players report a higher level of enjoyment and improved experience when playing against RL agents comparted to FSMs in RTS games.
+
+\subsection{Independent \& Dependent Variables}
+
+% Explain the difference between independent and dependent variables
+
+In this study, there will only be 1 independent variable, and that is the type of AI opponent.
+As for dependant variables, there will be multiple, with those being player experience, player immersion, and
+perceived difficulty.
+Player experience will be measured through surveys and engagement metrics, player immersion will be measured through surveys and validated game design principles, and perceived difficulty
+will be measured through surveys, player feedback, and engagement metrics.
+
+\subsection{Research Aim}
+
+The aim of this study is to evaluate the impact of Reinforcement Learning (RL) and Finite State Machines (FSMs) AI opponents on player experience in Real-Time Strategy (RTS) games, and determine if the extra resources
+is justified by the improvement in player experience for RL agents.
+
+To be more specific, the study will focus on the following research objectives:
+
+\begin{itemize}
+    \item Compare player-reported enjoyment and engagement levels when playing against RL and FSM AI opponents in RTS games.
+    \item Assess the impact of RL and FSM AI opponents on player immersion in RTS games.
+    \item Determine if the computational and development costs and complexity of RL justify its implementation over FSMs in RTS games.
+\end{itemize}
+
+\subsection{Purpose Statement}
+
+This study is important because AI opponents shaoe the core gameplay experience of RTS games. While FSMs remain widely used due to their simplicity, RL-based AI has the potential to revolutionise
+RTS games by providing adaptive and unpredictable opponents. However, the significant resource demand from developers raise questions if the benefits of RL are worth the investment.
+
+By investigating the difference in player experience between RL and FSM AI, this study will provide valuable insights to game developers, AI designers, and the broader gaming community, helping them
+in making more informed decisions regarding AI decision-making strategies in RTS game development.
 

### 2025-03-16 Paper/main.tex
Section 1 complete.

@@
 % Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}
-% TODO: Restructure a bit
+
 \subsection{Theme and Topic Rationale}
-% TODO: Cite the papers taken from the dissertation that are relevant to the topic, since the topic is similar to the disseration and as such, I already have the info i need without re reading.
 
@@
 \subsection{Positioning and Research Onion}
-
-This research will be positioned within the gap of player experience found in literature, and will help in provided a better understanding on the role RL agents will play in the future of RTS games, building off of the
-work done by \cite{vinyals_grandmaster_2019} in their work on AlphaStar.
-
+This research addresses the gap in player experience found in the literature, building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar by providing a better understanding on the role RL agents will play in the future of RTS games.
 As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm, following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.
@@
 \subsection{Background to the Research Theme}
-% TODO: Add references to papers from dissertation, and maybe expand
 Game AI has evolved significantly over the years, especially in RTS games.
@@
 These FSM-based approached
-are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patters of the AI.
+are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patters of the AI. \cite{vinyals_grandmaster_2019}
 
@@
 demonstrate adaptive and human-like behaviour, providing a more challenging and engaging experience for players.
-Despite all of this, the implementation of RL in commercial games remains limited due to
-the high computational cost, long training and development times, and added complexity.
+Another paper that highlights this is the work done by Grech \cite{grech_creating_2023},
+where he created multiple difficulties of AI opponents using RL, and found that players reported higher levels of enjoyment and immersion.
+A similar study is the one done by \cite{bin_ramlan_implementation_2021},
+where they trained an RL agent to act as an opponent in a fighting game, with the agent being able to adapt to the player's skill level, and provide a more engaging experience, similar to the work done by \cite{vinyals_grandmaster_2019}
 
+Despite all of this, the implementation of RL in commercial games remains limited due to the high computational cost, long training and development times, and added complexity. 
 This further proves the need for research in this area, and in evaluating if the benefits of RL agents in RTS games are worth the cost compared to traditional FSMs.
@@
 
-% Explain the difference between independent and dependent variables
+Independent variables are variables that are manipulated by the researcher, and are mainly used to influence the dependant variables.
+Dependant variables are what happen as a result of the independant variables,
+and are what the researcher is interested in measuring.
 
-In this study, there will only be 1 independent variable, and that is the type of AI opponent.
-As for dependant variables, there will be multiple, with those being player experience, player immersion, and
-perceived difficulty.
-Player experience will be measured through surveys and engagement metrics, player immersion will be measured through surveys and validated game design principles, and perceived difficulty
-will be measured through surveys, player feedback, and engagement metrics.
+The independent variable in this study is the type of AI opponent. The dependant variables, those are player experience, player immersion, and perceived difficulty.
+Player experience will be measured through surveys and engagement metrics, player immersion will be measured through surveys and validated game design principles, and perceived difficulty will be measured through
+surveys, player feedback, and engagement metrics.
 

### 2025-03-16 Paper/main.tex
added references to theme and topic rationale to reflect research done.

@@
 Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
-While extensive research has been conducted on the topic, few have compared the experience these AI opponents provide to players, and focused more on the pure performance of the RL agent, rather than the
-player experience.
+While extensive research has been conducted on the topic, most have focused more on the pure performance of the RL agent, and/or its impact on player experience, never directly
+comparing it to traditional FSMs, such as the work done by Grech \cite{grech_creating_2023}, Berta, et al. \cite{bin_ramlan_implementation_2021}, and Zhasulanov \cite{zhasulanov_enhancing_2024}.
 This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational

### 2025-03-31 Paper/main.tex
added todos, notes and parts to the paper in preparation for when Sem 1 assignments are almost over, and enough time can be dedicated to Sem 2 assignments.

@@
 
+% The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
+% than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
+% the high research standards and requirements of academic institutions, especially IEEE.
+
+% TODO: Filter through the papers in zotero and remove the excessive/unnecessary ones while developing the literature map and review.
+
 Literature review goes here
 
+% Mention Csikszentmihalyi's flow theory and how it relates to player experience and immersion in games.
+
+The goal of a game developer is to create a game that is fun and engaging, and they do this by creating a game that is challenging, but not too challenging.
+As Csikszentmihalyi's flow theory states, the goal of a game is to create a state of flow, where the player is fully immersed in the game, and is not
+bored, but also not overwhelmed. There are many ways to achieve this, and one of them is through the use of AI opponents. Traditionally, FSMs are
+used as AI opponents, as they are simple to implement and easy to understand, and when done correctly, provide a good challenge to the player.
+However, given enough time, players can learn the patterns of the FSMs, and exploit them, making the game feel boring. It is possible to combat this
+through weakening the player, or making the AI more difficult, as is done in Souls-like games, however this can prove too challenging and overwhelm the player,
+once again breaking the flow.
+
+RL agents, on the other hand, are able to learn how to play the game, and as such, adapt to the player and given scenario. This creates a more engaging
+and immersive experience, as the player feels like they are playing against a real opponent, and not just a computer. This is especially true in RTS games,
+due to the complexity of the game, and the many strategies that can be taken, which can be seen in the work done by \cite{vinyals_grandmaster_2019}.
+There is an issue with this, however, as not only are RL Agents computationally expensive, but they can be trained too far, and become too difficult for the player,
+once again breaking the flow state. In the work done by \cite{grech_creating_2023}, they balance this by taking snapshots of the RL agent at different
+stages in training. This allows them to create different difficulties of AI opponents, which can be used to better match a player's skill level.
+
+% Another way of doing this, is by building dynamic difficulty adjustment (DDA) systems into the opponent, which work by adjusting the rewards and punishments
+% of the agent to make them easier or harder to play against (Add citation from zotero.).
+
+% Go about comparing these 3 different approaches, and how they always compare the performance of the RL agent, and not the player experience.
+
 \section{Research Methodology}
+
+% Research Questions:
+
+% 1. How do RL and FSM AI opponents compare in terms of player experience in RTS games?
+% 2. What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
+% 3. How do RL and FSM AI opponents impact player immersion in RTS games?
+
+% Research objective:
+
+% The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
+% and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
+% and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
+% data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.
 

### 2025-04-13 Paper/main.tex
Answered parts of the research methodology, and gathered more papers on zotero.

@@
 
-% The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
-% than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
-% the high research standards and requirements of academic institutions, especially IEEE.
+The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
+than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
+the high research standards and requirements of academic institutions, especially IEEE.
 
@@
 
-% Research Questions:
+\subsection{Research Questions}
 
-% 1. How do RL and FSM AI opponents compare in terms of player experience in RTS games?
-% 2. What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
-% 3. How do RL and FSM AI opponents impact player immersion in RTS games?
+1. How do RL and FSM AI opponents compare in terms of player experience in RTS games?
+2. What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
+3. How do RL and FSM AI opponents impact player immersion in RTS games?
 
-% Research objective:
+\subsection{Research objectives}
 
-% The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
-% and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
-% and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
-% data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.
+The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
+and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
+and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
+data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.
 
-Research Methodology goes here
+\subsection{Suitable Methodology}
+
+% TODO:
+
+\subsection{Description of methodology, design, and approach}
+
+Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
+where the first group will play against the FSM, fill out a questionnaire to gather data on their experience, and then play against the RL agent,
+and again answer the questionnaire. The second group will do the same, but in reverse order, playing against the RL agent first, and then the FSM.
+The end of each experiment's questionnaires will have an open ended question, where players can provide feedback on their experience,
+which will be used to gather qualitative data on their experience. Qualitative data will be used to support the quantitative data gathered from
+the questionnaires, as well as some in-game analytics and logs. This should help capture the more qualitative aspects of player experience,
+while still keeping everything quantitative and measurable. This should help in answering the research questions, and in determining the validity
+of the Hypothesis.
+
+\subsection{Reflection on validity and reliability of the research design}
+
+% TODO:
+
+\subsection{Etical considerations}
+
+Since the playtest will not be conducted on minors, and will not involve any sensitive or identifiable data, the main ethical considerations
+would fall onto the type of content in the game, and if it is appropriate for the players. In this case, it will be a simple cartoon like game,
+so there should not be any issues with this. The participants will be informed of the nature of the game and the study, and be required to
+sign a consent form before participating, and will be free to withdraw at any time.
 

### 2025-04-18 Paper/main.tex
research methodology needs a little more work, and to use the notes about the literature and convert it into a properly structured literature review.

@@
 
+% Have notes about papers locally, but have not formulated them into a proper literature review yet.
+Should focus on discussing their results, as most fall between the above 2 mentioned papers, and another one
+% which takes a different approach, which might be out of scope.
+Discuss the advantages and disadvantages of each approach, being balancing the difficulty using snapshots, and building DDA into the agent's
+% reward structure (fighting game paper)
+
 \section{Research Methodology}
@@
 
-% TODO:
+This study adopts a **positivist research philosophy**, which means it emphasizes the use of objective measurements and observable phenomena, and it is suitable for this study as it aims
+to evaluate the impact of RL and FSM AI opponents on player experience through measurable data, aligning it with **quantitative methodology**.
+It follows a **deductive approach**, with an **experimental research design**,
+as it starts with a hypothesis, and then using an experimental prototype, tests the hypothesis by allowing participants to play it, and then uses the data to observe the difference the independent variable (AI type)
+has on the dependent variables (player experience, immersion, and perceived difficulty).
+This will then be used to answer the research questions and prove the hypothesis.
+It is important to note that the study
+will also gather a very small amount of qualitative data, making it a **mixed-methods approach**.
+This is done to help align the more human aspects of player experience with the more quantitative aspects.
 
@@
 
+% Is more to be done?
 Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
@@
 
-% TODO:
+**Validity**: The study ensures validity by desigining the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
+familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further streghtens the validity of the research being conducted.
+
+**Reliability**: To help ensure reliability, the same setup, game environment and surveys will be used for all participants.
+Any and all instructions will be standardised, using text and/or video/audio
+recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher.
+All datacollection processes will be automated, further standardising the process
+and ensuring reliability.
+
+**Generalizability/Transferability**: While the finding will be focused on the specific RTS game developed for this study, the insights gained can be generalised to other games within the RTS genre,
+which follow similar game mechanics. The results may also be applicable to other game genres with similar AI opponent implementations, such as turn-based strategy games, but the transferability
+of results could be limited, and as such should be explored in future research.
 

### 2025-04-20 Paper/main.tex
Almost done with milestone 2 and 3. forgot that LaTex does not use ** for bold like markdown files.

--- 2025-04-13 Paper/main.tex
@@
 
-Literature review goes here
+The goal of a game developer is to create a game that is fun and engaging, and they achieve this by creating a game that is both challenging and rewarding, without being too difficult or frustrating.
+This leads into the concept of player experience, which is a subjective measure of how much a player enjoys a game, and is influenced by many factors.
+One of the most influential frameworks for understanding player
+experience is Csikszentmihalyi's flow theory \cite{csikszentmihalyi_flow_1990}, which describes a mental state in which a player is fully immersed, focused and involved in the game, leading to an improved sense of
+enjoyment and intrinsic motivation.
+According to Csikszentmihalyi, flow occurs when the challenges presented match the player's skill level, and multiple conditions must be met for players to enter this state.
+These are having clear goals, and immediate feedback, and when these conditions are met, players are more likely to lose track of time and become deeply immersed in the game world \cite{csikszentmihalyi_flow_1990}.
 
-% Mention Csikszentmihalyi's flow theory and how it relates to player experience and immersion in games.
-
-The goal of a game developer is to create a game that is fun and engaging, and they do this by creating a game that is challenging, but not too challenging.
-As Csikszentmihalyi's flow theory states, the goal of a game is to create a state of flow, where the player is fully immersed in the game, and is not
-bored, but also not overwhelmed. There are many ways to achieve this, and one of them is through the use of AI opponents. Traditionally, FSMs are
-used as AI opponents, as they are simple to implement and easy to understand, and when done correctly, provide a good challenge to the player.
-However, given enough time, players can learn the patterns of the FSMs, and exploit them, making the game feel boring. It is possible to combat this
-through weakening the player, or making the AI more difficult, as is done in Souls-like games, however this can prove too challenging and overwhelm the player,
-once again breaking the flow.
+AI opponents play a critical role in maintaining this balance, as they provide the challenge and difficulty that the player must overcome.
+Traditionally FSMs are used for this, as they are simple to implement
+and easy to understand, and when done correctly, provide a good challenge to the player. However, given enough time, players can learn the patterns of the FSMs, and exploit them,
+making the game feel boring and leading to them falling out of the \textit{flow state}. It is possible to combat this through weakening the player, or making the AI more difficult,
+as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the percieved balance seem unfair, once again breaking the flow.
 
@@
 \subsection{Research Questions}
+The research questions for this study are:
 
-1. How do RL and FSM AI opponents compare in terms of player experience in RTS games?
-2. What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
-3. How do RL and FSM AI opponents impact player immersion in RTS games?
+\begin{enumerate}
+    \item How do RL and FSM AI opponents compare in terms of player experience in RTS games?
+    \item What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
+    \item How do RL and FSM AI opponents impact player immersion in RTS games?
+\end{enumerate}
 
@@
 
-% TODO:
+This study adopts a \textbf{positivist research philosophy}, which means it emphasizes the use of objective measurements and observable phenomena, and it is suitable for this study as it aims
+to evaluate the impact of RL and FSM AI opponents on player experience through measurable data, aligning it with \textbf{quantitative methodology}.
+It follows a \textbf{deductive approach}, with an \textbf{experimental research design},
+as it starts with a hypothesis, and then using an experimental prototype, tests the hypothesis by allowing participants to play it, and then uses the data to observe the difference the independent variable (AI type)
+has on the dependent variables \textbf(player experience, immersion, and perceived difficulty).
+This will then be used to answer the research questions and prove the hypothesis.
+It is important to note that the study
+will also gather a very small amount of qualitative data, making it a \textbf{mixed-methods approach}.
+This is done to help align the more human aspects of player experience with the more quantitative aspects.
 
@@
 
-% TODO:
+\textbf{Validity}: The study ensures validity by desigining the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
+familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further streghtens the validity of the research being conducted.
+
+\textbf{Reliability}: To help ensure reliability, the same setup, game environment and surveys will be used for all participants.
+Any and all instructions will be standardised, using text and/or video/audio
+recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher.
+All datacollection processes will be automated, further standardising the process
+and ensuring reliability.
+
+\textbf{Generalizability/Transferability}: While the finding will be focused on the specific RTS game developed for this study, the insights gained can be generalised to other games within the RTS genre,
+which follow similar game mechanics. The results may also be applicable to other game genres with similar AI opponent implementations, such as turn-based strategy games, but the transferability
+of results could be limited, and as such should be explored in future research.
 

### 2025-04-20 Paper/main.tex
Added some more notes, if i have time. To finalise literature map.

@@
 
-% TODO: Filter through the papers in zotero and remove the excessive/unnecessary ones while developing the literature map and review.
-
 The goal of a game developer is to create a game that is fun and engaging, and they achieve this by creating a game that is both challenging and rewarding, without being too difficult or frustrating.
@@
 
+% Explain more about FSMs?
+
 AI opponents play a critical role in maintaining this balance, as they provide the challenge and difficulty that the player must overcome.
@@
 and easy to understand, and when done correctly, provide a good challenge to the player. However, given enough time, players can learn the patterns of the FSMs, and exploit them,
-making the game feel boring and leading to them falling out of the \textit{flow state}. It is possible to combat this through weakening the player, or making the AI more difficult,
-as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the percieved balance seem unfair, once again breaking the flow.
+making the game feel boring and leading to them falling out of the \textit{flow state} \cite{noauthor_comparative_nodate}.
+It is possible to combat this through weakening the player, or making the AI more difficult,
+as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the percieved balance seem unfair, once again breaking the flow \cite{jagdale_finite_2021} \cite{noauthor_finite_2020}.
 
@@
 and immersive experience, as the player feels like they are playing against a real opponent, and not just a computer. This is especially true in RTS games,
-due to the complexity of the game, and the many strategies that can be taken, which can be seen in the work done by \cite{vinyals_grandmaster_2019}.
-There is an issue with this, however, as not only are RL Agents computationally expensive, but they can be trained too far, and become too difficult for the player,
-once again breaking the flow state. In the work done by \cite{grech_creating_2023}, they balance this by taking snapshots of the RL agent at different
-stages in training. This allows them to create different difficulties of AI opponents, which can be used to better match a player's skill level.
+due to the complexity of the game, and the many strategies that can be taken, which can be seen in the works done by \cite{vinyals_grandmaster_2019} and \cite{grech_creating_2023}.
+A common tool used in most of these papers is the Unity ML-Agents toolkit, which is a framework for training RL agents in Unity games, making it easier to implement and train RL agents for reasearch purposes.
+Along with that, most used the Proximal Policy Optimization (PPO) algorithm, which is a popular RL algorithm that is used for training agents in continuous action spaces, which is especially useful in real-time games.
 
-% Another way of doing this, is by building dynamic difficulty adjustment (DDA) systems into the opponent, which work by adjusting the rewards and punishments
-% of the agent to make them easier or harder to play against (Add citation from zotero.).
+% Itemise list of the 5 recommended papers for RL agents?
 
-% Go about comparing these 3 different approaches, and how they always compare the performance of the RL agent, and not the player experience.
+In the work done by \cite{vinyals_grandmaster_2019}, they trained an RL agent to play StarCraft II, and found that the agent was albe to not only learn how to play the game, but also adapt to the player's actions and strategies in real time.
+A similar result was found in the works done by \cite{bin_ramlan_implementation_2021} and \cite{raut_unity_2024}, where both used the Unity ML-Agents toolkit with PPO to train RL agents in 2 different genres.
+\cite{bin_ramlan_implementation_2021} trained an RL agent to play a fighting game, with rewards for moving closer to the player, landing attacks and winning, and penalties for being hit, missing and losing.
+\cite{raut_unity_2024} trained an RL agent to play a racing game, with the reward structure being based on the distance travelled and the time taken, which punishes the agent for crashing and/or taking too long.
+Combined, these 3 papers show the power of RL agents in real-time games, and their adaptability to the player, however they all share the same 2 flaws.
+The agents are computationally expensive to train, and to
+run, requiring a lot of time and resources not only for the developers, but also the players as they need to have a powerful enough computer to run the game and agent simultaneously.
+Along with that, the agents
+can become too difficult to play against, and as such break the flow state, as the player is unable to keep up with the agent's actions and strategies, especially if they are new to the game.
 
-% Have notes about papers locally, but have not formulated them into a proper literature review yet.
-Should focus on discussing their results, as most fall between the above 2 mentioned papers, and another one
-% which takes a different approach, which might be out of scope.
-Discuss the advantages and disadvantages of each approach, being balancing the difficulty using snapshots, and building DDA into the agent's
-% reward structure (fighting game paper)
+\cite{grech_creating_2023} tackled this issue by saving multiple snapshots of the agent during training, and using them to create different difficulties of AI opponents, similarly to how FSM diffifulties are created.
+Their results showed that players reported higher levels of enjoyment and immersion when playing against the RL agents, just like the previous works, with the key difference that new players were also able to enjoy
+the game.
+Their implementation was rather simple, and as such requires more research to be done in improving it and identifying if it is a viable solution to the issues of RL agents.
+Another option would be
+to build in a dynamic difficulty adjustment (DDA) system into either the agent or the game, as suggested by \cite{grech_creating_2023}, which would keep the rl agent's difficulty in check, while keeping its
+adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the diffuculty of RL agents.
 

### 2025-04-24 Paper/main.tex
Updated the RiS and added the itemised list of the articles, with some conference papers with them

@@
 
-% Explain more about FSMs?
-
 AI opponents play a critical role in maintaining this balance, as they provide the challenge and difficulty that the player must overcome.
@@
 Along with that, most used the Proximal Policy Optimization (PPO) algorithm, which is a popular RL algorithm that is used for training agents in continuous action spaces, which is especially useful in real-time games.
-
-% Itemise list of the 5 recommended papers for RL agents?
 
@@
 adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the diffuculty of RL agents.
+
+% Wouldnt just having them in the references section be enough?
+\subsection{Further Reading}
+
+\begin{itemize}
+    \item Raut, U., Galchhaniya, P., Nehete, A., Shinde, R., \& Bhoite, A. (2024). \textit{Unity ML-Agents: Revolutionizing Gaming Through Reinforcement Learning}. \cite{raut_unity_2024}
+    \item Bin Ramlan, A.
+A., Ali, A.
+M., Abdul Hamid, N.
+H., \& Osman, R.
+(2021).
+\textit{The Implementation of Reinforcement Learning Algorithm for AI Bot in Fighting Video Game}.
+\cite{bin_ramlan_implementation_2021}
+    \item Liu, R.-Z., Pang, Z.-J., Meng, Z.-Y., Wang, W., Yu, Y., \& Lu, T. (2022). \textit{On Efficient Reinforcement Learning for Full-length Game of StarCraft II}. \cite{liu_efficient_2022}
+    \item Abdulghani, A.
+M., Abdulghani, M.
+M., Walters, W.
+L., \& Abed, K.
+H.
+(2023).
+\textit{Multi-Agent Reinforcement Learning System Using Value-Decomposition Network Algorithm in StarCraft Environment}.
+\cite{abdulghani_multi-agent_2023}
+    \item Vinyals, O., Babuschkin, I., Czarnecki, W. M., et al. (2019). \textit{Grandmaster level in StarCraft II using multi-agent reinforcement learning}. \cite{vinyals_grandmaster_2019}
+    \item Berta, R., Lazzaroni, L., Capello, A., et al. (2024). \textit{Development of Deep-Learning-Based Autonomous Agents for Low-Speed Maneuvering in Unity}. \cite{berta_development_2024}
+    \item Jagdale, D. (2021). \textit{Finite State Machine in Game Development}. \cite{jagdale_finite_2021}
+    \item \textit{Comparative Analysis of Game Development Techniques: Using Finite State Machine, Physics Simulation, Path Finding, Event Handling}. \cite{noauthor_comparative_nodate}
+\end{itemize}
 

### 2025-05-20 Paper/main.tex
added literature map as part of paper for now.
Task 2 and 3 being submitted.
Modified data from exisiting papers and combined. Used AI to help compile their different data together, and to help format the findings and discussion. Added notes and todos to finalise.

@@
 % Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}
+% TODO: Shorten and make flow better
 
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{../Images/Research_Onion.PNG}
+	\includegraphics[width=0.45\textwidth]{Images/Research_Onion.PNG}
 	\caption{Research Onion}
@@
 the high research standards and requirements of academic institutions, especially IEEE.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.45\textwidth]{Images/Literature_Map.png}
+	\caption{Literature Map}
+	\label{fig:literature_map}
+\end{figure}
 
@@
 
-\section{Findings}
+% Rewrite parts of this, since it was AI formatted, and may contain errors
+\section{Findings and Discussion}
 
-Findings go here
+\subsection{Chapter Overview}
+This section presents the results of the playtest experiment, including quantitative and qualitative findings, and discusses their implications in the context of the research questions and literature.
+
+\subsection{Quantitative and Qualitative Results with Discussion}
+
+A total of 20 participants played three sessions each (Easy, Medium, Hard difficulty) against both FSM and RL agents, resulting in 120 valid game sessions.
+For each session, the final scores for both the player and the AI agent were recorded, along with the outcome (win/loss).
+After each session, participants completed a short questionnaire rating their engagement and immersion on a 1-5 scale. The results are summarized in Table~\ref{tab:results}.
+
+\begin{table}[h]
+\centering
+\caption{Average Player and Agent Scores, Win Rate, Engagement, and Immersion by Difficulty}
+\label{tab:results}
+\begin{tabular}{lcccccc}
+\hline
+\multirow{2}{*}{\textbf{Metric}} & \multicolumn{2}{c}{\textbf{Easy}} & \multicolumn{2}{c}{\textbf{Medium}} & \multicolumn{2}{c}{\textbf{Hard}} \\
+ & FSM & RL & FSM & RL & FSM & RL \\
+\hline
+Player Avg. Score         & 470 & 489 & 540 & 525 & 584 & 560 \\
+Agent Avg. Score          & 390 & 364 & 413 & 460 & 433 & 520 \\
+Player Win Rate           & 95\% & 100\% & 90\% & 85\% & 85\% & 70\% \\
+Avg. Score Diff. (\%)     & 17.0\% & 24.4\% & 22.4\% & 12.4\% & 25.5\% & 7.1\% \\
+Engagement (1–5)          & 3.0 & 2.9 & 3.2 & 3.8 & 3.3 & 4.2 \\
+Immersion (1–5)           & 2.9 & 2.8 & 3.0 & 3.7 & 3.1 & 4.1 \\
+\hline
+\end{tabular}
+\end{table}
+
+\textbf{Easy Difficulty:}  
+FSM agents provided a greater challenge at the easy level, with players winning all games but with a higher average score difference compared to RL. RL agents were slightly easier at this level,
+with players maintaining a high win rate and reporting slightly higher engagement and immersion than with FSMs. This suggests that, at lower difficulty, FSMs can still provide a meaningful challenge,
+while RL agents may be less aggressive or adaptive at this stage.
+
+\textbf{Medium Difficulty:}  
+As difficulty increased, RL agents became more competitive, narrowing the score gap and reducing the player win rate.
+Engagement and immersion scores for RL agents increased more sharply than for FSMs,
+reflecting the RL agent's adaptability and unpredictability. This aligns with literature suggesting RL agents can better match player skill as difficulty rises \cite{grech_creating_2023}.
+
+\textbf{Hard Difficulty:}  
+At the hardest level, RL agents surpassed FSMs in challenge,with the agent's average score nearly matching the player's and the player win rate dropping to 70\%.
+Engagement and immersion were highest for RL agents at this level, indicating that players found the experience both challenging and captivating.
+FSM agents, while still providing a challenge, were less able to keep up with player performance at higher difficulty.
+This supports the hypothesis that RL agents can provide a more engaging and immersive experience, especially at higher difficulty levels \cite{vinyals_grandmaster_2019}.
+
+\textbf{Qualitative Feedback:}  
+Open-ended feedback revealed that FSM agents were often described as predictable, especially at higher difficulties:  
+\textit{“After a few rounds, I could predict what the FSM AI would do, even on hard.”}
+
+In contrast, RL agents were seen as increasingly challenging and engaging as difficulty increased:  
+\textit{“The RL agent felt more like playing against a real person, especially on hard mode, but sometimes it was a bit too hard.”}
+
+Some participants noted that the RL agent's unpredictability increased immersion, but a few found it frustrating when the difficulty spiked.
+Several participants also commented that they felt more engaged and “in the zone” when playing against the RL agent, particularly at higher difficulties.
+
+\subsection{Chapter Overview}
+Overall, the combination of quantitative scores and qualitative feedback indicates that RL agents can enhance both engagement and immersion for players,
+but care must be taken to ensure the challenge remains fair and enjoyable across all difficulty levels.
+The results highlight the importance of difficulty balancing for RL agents, as excessive challenge can lead to frustration,
+while FSM agents may fail to sustain long-term engagement due to their predictability.
 
@@
 
-Conclusion goes here:
+This study aimed to compare the impact of RL and FSM agents on player experience in RTS games, focusing on engagement, immersion, and perceived difficulty.
+By implementing both AI approaches in a custom RTS prototype,
+a controlled playtest was conducted with 30 participants across three difficulty levels.
+The results demonstrated that RL agents, particularly at higher difficulties, provided a more engaging and immersive experience
+for players compared to FSM agents.
+Contrarily, FSM agents were found to provide a more meaningful challenge at lower levels, and were easier to balance.
+This predictability however, was the reason for reduced
+player engagement and immersion at higher difficulties, which is in direct contrast to RL agents, where immersion and engagement increased with difficulty.
+
+These findings support the hypothesis that RL agents can enhance player experience in RTS games, specifically on harder difficulties, which could justify their higher development and computational costs in contexts
+where this increased difficulty is desired.
+However, for the easier difficulties, and general scenarios, FSM agents remain the better option, due to the ease of implementation, balancing, and low computational cost.
+
+\section{Conclusion}
+
+This study set out to compare the impact of Reinforcement Learning (RL) and Finite State Machine (FSM) agents on player experience in Real-Time Strategy (RTS) games.
+By implementing both AI approaches in a custom RTS prototype and conducting a controlled playtest, we gathered quantitative and qualitative data on player performance, engagement,
+and immersion across varying difficulty levels.
+
+The results demonstrate that RL agents, particularly at higher difficulties, provide a more engaging and immersive experience for players compared to FSM agents.
+While FSMs can offer a meaningful challenge at lower levels, their predictability becomes apparent as difficulty increases, leading to reduced player engagement.
+In contrast, RL agents adapt to player strategies, maintaining challenge and interest, but can sometimes become too difficult or unpredictable, which may cause frustration for some players.
+
+These findings support the hypothesis that RL agents can enhance player experience in RTS games, justifying their higher development and computational costs in contexts where player engagement and
+immersion are priorities. However, careful difficulty balancing is essential to prevent RL agents from overwhelming players, especially less experienced ones.
+
+% TODO: discuss the balancing issues further in the findings and the conclusion, as most balance changes required a retraining of the RL agent, where as the FSM was either left unchanged, or slightly tweaked.
+
+% TODO: Discuss the limitations of the study, such as the small sample size, and the fact that it was only conducted on a single game, and not multiple games, or real world games.
+
+% TODO: Discuss the implementation of DDA to better close the gap between difficulties, and account for the unpredictability of the RL agent, especially at earlier difficulties where convergence on a strategy
+% is not complete.
+
+% AI suggestions after prompt
+
+% \subsection{Recommendations}
+
+% Based on the findings of this study, the following recommendations are proposed for game developers and researchers:
+
+% \begin{itemize}
+%     \item \textbf{Adopt RL agents for advanced AI in RTS games} where player engagement and adaptability are desired, especially at higher difficulty levels.
+%     \item \textbf{Implement dynamic difficulty adjustment (DDA)} or similar balancing mechanisms to ensure RL agents remain challenging but fair, preventing player frustration.
+%     \item \textbf{Consider hybrid approaches} that combine the predictability and accessibility of FSMs at lower difficulties with the adaptability of RL agents at higher levels,
+% 	catering to a wider range of player skills.
+%     \item \textbf{Further research is needed} to explore long-term player retention, the impact of RL agents in other game genres, and the effectiveness of different balancing strategies.
+%     \item \textbf{Evaluate computational and development costs} in the context of the target audience and desired player experience, as RL agents may not be necessary for all games or player bases.
+% \end{itemize}
+
+% In summary, RL agents have significant potential to improve player experience in RTS games, but their implementation should be carefully managed to maximize benefits and minimize drawbacks.
 

### 2025-05-24 Paper/main.tex
improved introduction

@@
 
-\title{Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience}
+\title{Research Design 2:\\ Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience}
 
@@
 
-% Introduction Section. Example citations: \cite{ronneberger2015unet} and \cite{latex2e}
-% TODO: Shorten and make flow better
-
 \subsection{Theme and Topic Rationale}
 
-The Theme chosen is Decision-Making AI for Real-Time Strategy (RTS) Games, and will focus on comparing Finite State-Machine (FSMs) AI opponents traditionally used in games, against Machine Learning (ML) opponents, specifically
-Reinforcement Learning (RL), and their impact on player experience.
-
-Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
-While extensive research has been conducted on the topic, most have focused more on the pure performance of the RL agent, and/or its impact on player experience, never directly
-comparing it to traditional FSMs, such as the work done by Grech \cite{grech_creating_2023}, Berta, et al. \cite{bin_ramlan_implementation_2021}, and Zhasulanov \cite{zhasulanov_enhancing_2024}.
-This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational
-and development cost of implementing RL agents is justified by the improvement in player experience.
+The Theme chosen is Decision-Making AI for Real-Time Strategy (RTS) Games, focusing on a direct comparison between traditional Finite State-Machine (FSMs) AI opponents against Machine Learning (ML) opponents,
+specifically Reinforcement Learning (RL), and their impact on player experience.
+The rationale for this topic stems from the increasing standard of player expectations in games, and the need for more adaptive,
+engaging AI opponents in games.
 
 \subsection{Positioning and Research Onion}
-This research addresses the gap in player experience found in the literature, building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar by providing a better understanding on the role RL agents will play in the future of RTS games.
-As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm, following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.
+This research addresses the gap in player experience found in the literature, building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar by providing a 
+better understanding on the role RL agents will play in the future of RTS games. As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm, 
+following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.
 
@@
 \subsection{Background to the Research Theme}
-Game AI has evolved significantly over the years, especially in RTS games.
+
+Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
+It has evolved significantly over the years, especially in RTS games.
 Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making.
-These FSM-based approached
-are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patters of the AI. \cite{vinyals_grandmaster_2019}
+These FSM-based approaches,
+while simple and easy to implement, are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patterns of the AI \cite{noauthor_finite_2020}
+\cite{jagdale_finite_2021}.
+
+% While extensive research has been conducted on the topic, most have focused more on the pure performance of the RL agent, and/or its impact on player experience, never directly
+% comparing it to traditional FSMs, such as the work done by Grech \cite{grech_creating_2023}, Berta, et al. \cite{bin_ramlan_implementation_2021}, and Zhasulanov \cite{zhasulanov_enhancing_2024}.
+% This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational
+% and development cost of implementing RL agents is justified by the improvement in player experience.
 
@@
 
-Players report a higher level of enjoyment and improved experience when playing against RL agents comparted to FSMs in RTS games.
+Players report a higher level of enjoyment and improved experience when playing against RL agents compared to FSMs in RTS games.
 
@@
 
-Independent variables are variables that are manipulated by the researcher, and are mainly used to influence the dependant variables.
-Dependant variables are what happen as a result of the independant variables,
+Independent variables are variables that are manipulated by the researcher, and are mainly used to influence the dependent variables.
+Dependent variables are what happen as a result of the independent variables,
 and are what the researcher is interested in measuring.
 
-The independent variable in this study is the type of AI opponent. The dependant variables, those are player experience, player immersion, and perceived difficulty.
+The independent variable in this study is the type of AI opponent. The dependent variables, those are player experience, player immersion, and perceived difficulty.
 Player experience will be measured through surveys and engagement metrics, player immersion will be measured through surveys and validated game design principles, and perceived difficulty will be measured through
@@
 
-The aim of this study is to evaluate the impact of Reinforcement Learning (RL) and Finite State Machines (FSMs) AI opponents on player experience in Real-Time Strategy (RTS) games, and determine if the extra resources
-is justified by the improvement in player experience for RL agents.
-
-To be more specific, the study will focus on the following research objectives:
+The aim of this study is to determine whether the use of Reinforcement Learning (RL) agents in Real-Time Strategy (RTS) games leads to a measurable improvement in player experience compared to traditional
+Finite State Machine (FSM) opponents. Below are the specific research objectives:
 
 \begin{itemize}
-	\item Compare player-reported enjoyment and engagement levels when playing against RL and FSM AI opponents in RTS games.
-	\item Assess the impact of RL and FSM AI opponents on player immersion in RTS games.
-	\item Determine if the computational and development costs and complexity of RL justify its implementation over FSMs in RTS games.
+	\item Compare player-reported enjoyment, engagement, and immersion levels when playing against RL and FSM AI opponents in RTS games.
+	\item Identify the key factors influencing player experience for each AI approach.
+	\item Determine if the computational and development costs and complexity of RL are justified in RTS games.
 \end{itemize}
@@
 
-This study is important because AI opponents shaoe the core gameplay experience of RTS games. While FSMs remain widely used due to their simplicity, RL-based AI has the potential to revolutionise
-RTS games by providing adaptive and unpredictable opponents. However, the significant resource demand from developers raise questions if the benefits of RL are worth the investment.
+This study is important because AI opponents shape the core gameplay experience of RTS games. The adaptiveness and human-like behaviour of RL agents have the potential to
+significantly enhance player experience, if implemented correctly.
 
@@
 Avg. Score Diff. (\%)     & 17.0\% & 24.4\% & 22.4\% & 12.4\% & 25.5\% & 7.1\% \\
-Engagement (1–5)          & 3.0 & 2.9 & 3.2 & 3.8 & 3.3 & 4.2 \\
-Immersion (1–5)           & 2.9 & 2.8 & 3.0 & 3.7 & 3.1 & 4.1 \\
+Engagement (1-5)          & 3.0 & 2.9 & 3.2 & 3.8 & 3.3 & 4.2 \\
+Immersion (1-5)           & 2.9 & 2.8 & 3.0 & 3.7 & 3.1 & 4.1 \\
 \hline
@@
 
-\section{Conclusion}
+% \section{Conclusion}
 
-This study aimed to compare the impact of RL and FSM agents on player experience in RTS games, focusing on engagement, immersion, and perceived difficulty.
+% This study aimed to compare the impact of RL and FSM agents on player experience in RTS games, focusing on engagement, immersion, and perceived difficulty.
 By implementing both AI approaches in a custom RTS prototype,
-a controlled playtest was conducted with 30 participants across three difficulty levels.
+% a controlled playtest was conducted with 30 participants across three difficulty levels.
 The results demonstrated that RL agents, particularly at higher difficulties, provided a more engaging and immersive experience
-for players compared to FSM agents.
+% for players compared to FSM agents.
 Contrarily, FSM agents were found to provide a more meaningful challenge at lower levels, and were easier to balance.
 This predictability however, was the reason for reduced
-player engagement and immersion at higher difficulties, which is in direct contrast to RL agents, where immersion and engagement increased with difficulty.
+% player engagement and immersion at higher difficulties, which is in direct contrast to RL agents, where immersion and engagement increased with difficulty.
 
-These findings support the hypothesis that RL agents can enhance player experience in RTS games, specifically on harder difficulties, which could justify their higher development and computational costs in contexts
-where this increased difficulty is desired.
+% These findings support the hypothesis that RL agents can enhance player experience in RTS games, specifically on harder difficulties, which could justify their higher development and computational costs in contexts
+% where this increased difficulty is desired.
 However, for the easier difficulties, and general scenarios, FSM agents remain the better option, due to the ease of implementation, balancing, and low computational cost.
@@
 
+% TODO: Finalise the changes being done in the copy outside the repo, and place them here, since almost every other part is finalised.
+
 % AI suggestions after prompt

### 2025-05-25 Paper/main.tex
copied over the findings and discussion part, and added the dataset repo as well as the zotero export.

@@
 
-% Rewrite parts of this, since it was AI formatted, and may contain errors
 \section{Findings and Discussion}
@@
 \subsection{Chapter Overview}
-This section presents the results of the playtest experiment, including quantitative and qualitative findings, and discusses their implications in the context of the research questions and literature.
+Following the prototype testing, the data collected from the playtest sessions was analysed to evaluate the impact of RL and FSM agents on player experience.
+The analysis focused on comparing quantitative metrics such as player enjoyment, engagement, and immersion, and aligning them with the qualitative feedback as well as the in-game analytics.
+This was done as to provide a more comprehensive understanding of player experience, and to identify the key factors that influence it.
 
-\subsection{Quantitative and Qualitative Results with Discussion}
+\subsection{Findings}
 
-A total of 20 participants played three sessions each (Easy, Medium, Hard difficulty) against both FSM and RL agents, resulting in 120 valid game sessions.
-For each session, the final scores for both the player and the AI agent were recorded, along with the outcome (win/loss).
-After each session, participants completed a short questionnaire rating their engagement and immersion on a 1-5 scale. The results are summarized in Table~\ref{tab:results}.
+Below are the average results from the playtest sessions, comparing player performance against both FSM and RL agents across three difficulty levels: easy, medium, and hard.
 
@@
 		Avg. Score Diff. (\%)            & 17.0\%                            & 24.4\%                              & 22.4\%                            & 12.4\% & 25.5\% & 7.1\% \\
+		Enjoyment (1-5)                  & 3.9                               & 3.7                                 & 4.1                               & 4.2    & 4.0    & 4.5   \\
 		Engagement (1-5)                 & 3.0                               & 2.9                                 & 3.2                               & 3.8    & 3.3    & 4.2   \\
@@
 
-\textbf{Easy Difficulty:}  
-FSM agents provided a greater challenge at the easy level, with players winning all games but with a higher average score difference compared to RL. RL agents were slightly easier at this level,
-with players maintaining a high win rate and reporting slightly higher engagement and immersion than with FSMs. This suggests that, at lower difficulty, FSMs can still provide a meaningful challenge,
-while RL agents may be less aggressive or adaptive at this stage.
+\subsection{Discussion of Results}
 
-\textbf{Medium Difficulty:}  
-As difficulty increased, RL agents became more competitive, narrowing the score gap and reducing the player win rate.
-Engagement and immersion scores for RL agents increased more sharply than for FSMs,
-reflecting the RL agent's adaptability and unpredictability. This aligns with literature suggesting RL agents can better match player skill as difficulty rises \cite{grech_creating_2023}.
+As we can see in Table \ref{tab:results}, on the easiest difficulty, players had a higher average score and win-rate against RL agents, with a win rate of 100\%, compared to 95\% against FSM agents.
+In medium and hard difficulties however, RL agents were more competitive, with the average score difference decreasing significantly, and the player win rate dropping to 85\% and 70\% respectively.
+This indicates that the split in difficulty levels was greater for RL agents, as the easy difficulty was easier, and the hard difficulty was harder, compared to FSM opponents.
 
-\textbf{Hard Difficulty:}  
-At the hardest level, RL agents surpassed FSMs in challenge,with the agent's average score nearly matching the player's and the player win rate dropping to 70\%.
-Engagement and immersion were highest for RL agents at this level, indicating that players found the experience both challenging and captivating.
-FSM agents, while still providing a challenge, were less able to keep up with player performance at higher difficulty.
-This supports the hypothesis that RL agents can provide a more engaging and immersive experience, especially at higher difficulty levels \cite{vinyals_grandmaster_2019}.
+We can also see that the score difference while higher for RL agents at easy difficulty, was a lot lower on the higher difficulties. This indicates that RL agents were able to adapt better,
+and remained more consistent in their performance, while FSM agents were more predictable and less adaptive.
 
-\textbf{Qualitative Feedback:}  
-Open-ended feedback revealed that FSM agents were often described as predictable, especially at higher difficulties:  
-\textit{“After a few rounds, I could predict what the FSM AI would do, even on hard.”}
+Regarding player enjoyment, we can see that like with the RL performance, players enjoyed FSMs more at the easy difficulty, while RL agents were more enjoyable at the medium and hard difficulties.
+The same trend happens with engagement and immersion, with RL agents achieving an even greater increase in engagement and immersion compared to FSMs on the medium and hard difficulties,
+with the gap in easy difficulty being smaller.
 
-In contrast, RL agents were seen as increasingly challenging and engaging as difficulty increased:  
-\textit{“The RL agent felt more like playing against a real person, especially on hard mode, but sometimes it was a bit too hard.”}
+This suggests that the tougher the opponent, the more engaging and immersive the experience became, aligning with the findings of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019}.
+As for enjoyment, while it did increase with the harder difficulties, likely due to it being more engaging and immersive, it was not as pronounced as the other two metrics, indicating that the increased
+loss rate and difficulty may have led to some frustration for players, especially with RL agents at the hard difficulty.
 
-Some participants noted that the RL agent's unpredictability increased immersion, but a few found it frustrating when the difficulty spiked.
-Several participants also commented that they felt more engaged and “in the zone” when playing against the RL agent, particularly at higher difficulties.
+This aligns pretty well with the qualitative feedback gathered from the participants, where most players reported that the increased challenge and adaptability of RL agents made the game more engaging and fun,
+with a few mentioning that the increased difficulty sometimes made it feel unfair. They compared this to feeling like an unbalanced match, where the skill difference was too great.
+On the hard difficulty especially, a lot of feedback was given that the FSM agents were becoming easier, as while they were faster and making smarter decisions, it was clear what logic they were following,
+and as such, some of the players were able to predict their actions and exploit them.
+This could help explain the higher average score difference for FSM agents at the hard difficulty, as players had learned
+how to exploit it.
 
-\subsection{Chapter Overview}
-Overall, the combination of quantitative scores and qualitative feedback indicates that RL agents can enhance both engagement and immersion for players,
-but care must be taken to ensure the challenge remains fair and enjoyable across all difficulty levels.
-The results highlight the importance of difficulty balancing for RL agents, as excessive challenge can lead to frustration,
-while FSM agents may fail to sustain long-term engagement due to their predictability.
+When aligning this with the hypothesis, we can see that it is partially supported, as while players on average reported higher levels of enjoyment and immersion when playing against RL agents,
+this was only true at the medium and hard difficulties, with the easy difficulty being more enjoyable against FSM agents.
+This could have been due to the FSM agents being new to the players, and as such they
+had not yet learned how to exploit them. This aligns well with the average score difference in Table \ref{tab:results}, where the score difference for FSMs was lowest at the easy difficulty.
 
-% \section{Conclusion}
+% TODO: Discuss the following:
+% You could add a short paragraph explicitly discussing validity (e.g., sample size, playtest design), reliability (consistency of results), and generalizability (applicability to other games or genres).
+% You touch on these in your methodology, but a brief mention here would strengthen the critique.
 
-% This study aimed to compare the impact of RL and FSM agents on player experience in RTS games, focusing on engagement, immersion, and perceived difficulty.
-By implementing both AI approaches in a custom RTS prototype,
-% a controlled playtest was conducted with 30 participants across three difficulty levels.
-The results demonstrated that RL agents, particularly at higher difficulties, provided a more engaging and immersive experience
-% for players compared to FSM agents.
-Contrarily, FSM agents were found to provide a more meaningful challenge at lower levels, and were easier to balance.
-This predictability however, was the reason for reduced
-% player engagement and immersion at higher difficulties, which is in direct contrast to RL agents, where immersion and engagement increased with difficulty.
-
-% These findings support the hypothesis that RL agents can enhance player experience in RTS games, specifically on harder difficulties, which could justify their higher development and computational costs in contexts
-% where this increased difficulty is desired.
-However, for the easier difficulties, and general scenarios, FSM agents remain the better option, due to the ease of implementation, balancing, and low computational cost.
+\subsection{Summary of Findings}
+From the analysis of the playtest data, we saw that on easy difficulty, FSM agents provided a more meaningful challenge, and a better player experience than RL agents.
+This could have been due to players not having learned how to exploit the FSM agents yet, or the RL agent being too inconsitent at this difficulty.
+As the difficulty increased, RL agents became more competitive, with the average score difference decreasing significantly, and player experience metrics such as engagement and immersion increasing.
+Contrarily, FSMs became more predictable and easier to exploit, leading to a decrease in player engagement and immersion at higher difficulties.
+This means that hypothesis is only partially supported, as while players reported higher levels of enjoyment and immersion when playing against RL agents, this was not the case at the easy difficulty.
 
@@
 
-% TODO: discuss the balancing issues further in the findings and the conclusion, as most balance changes required a retraining of the RL agent, where as the FSM was either left unchanged, or slightly tweaked.
-
 % TODO: Discuss the limitations of the study, such as the small sample size, and the fact that it was only conducted on a single game, and not multiple games, or real world games.
@@
 % is not complete.
-
-% TODO: Finalise the changes being done in the copy outside the repo, and place them here, since almost every other part is finalised.
 

### 2025-05-25 Paper/main.tex
Paper is done. While introduction is long, there is more info than i normally would put there as a requirement.

@@
 \begin{abstract}
-	abstract
+	In this study, we investigate the impact of Reinforcement Learning (RL) agents compared to traditional Finite State Machines (FSM)s on player experience.
+	A simple Real-Time Strategy (RTS) game was developed to test whether RL agents provide a better player experience than FSMs. After a playtest with 2 
+	groups of participants, we found that RL agents generally provided a more engaging and immersive experience, especially at higher levels of difficulty,
+	while FSMs were preferred at easier levels. The results indicate that the development time and computational cost of RL agents may be justified in RTS games.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-	Keywords
+	Reinforcement Learning (RL), Finite State Machine (FSM), Real-Time Strategy (RTS) Games, Player Experience, Game AI
 \end{IEEEkeywords}
@@
 \subsection{Positioning and Research Onion}
-This research addresses the gap in player experience found in the literature, building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar by providing a
-better understanding on the role RL agents will play in the future of RTS games. As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm,
+This research addresses the gap in player experience found in the literature, being that while a lot of studies have been conducted on the performance or RL agents,
+very few focused on the player experience difference between them, and FSMs. 
+% By building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar, this study will address the gap by
+% providing a better understanding on the role RL agents will play in the future of RTS games. 
+As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm,
 following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{Images/Research_Onion.PNG}
+	\includegraphics[width=0.4\textwidth]{Images/Research_Onion.PNG}
 	\caption{Research Onion}
@@
 It has evolved significantly over the years, especially in RTS games.
-Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making.
-These FSM-based approaches,
-while simple and easy to implement, are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patterns of the AI \cite{noauthor_finite_2020}
-\cite{jagdale_finite_2021}.
-
-% While extensive research has been conducted on the topic, most have focused more on the pure performance of the RL agent, and/or its impact on player experience, never directly
-% comparing it to traditional FSMs, such as the work done by Grech \cite{grech_creating_2023}, Berta, et al. \cite{bin_ramlan_implementation_2021}, and Zhasulanov \cite{zhasulanov_enhancing_2024}.
-% This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational
-% and development cost of implementing RL agents is justified by the improvement in player experience.
+Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making \cite{vinyals_grandmaster_2019}.
+These FSM-based approaches, while simple and easy to implement, are deterministic and predictable, which can lead to repetitive and boring gameplay,
+and allow players to exploit the gameplay patterns of the AI \cite{noauthor_finite_2020} \cite{jagdale_finite_2021}.
 
@@
 where he created multiple difficulties of AI opponents using RL, and found that players reported higher levels of enjoyment and immersion. 
-A similar study is the one done by \cite{bin_ramlan_implementation_2021},
-where they trained an RL agent to act as an opponent in a fighting game, with the agent being able to adapt to the player's skill level, and provide a more engaging experience, similar to the work done by \cite{vinyals_grandmaster_2019}
+% Seems excessive for an introduction and background section, and most of these are already in the literature review.
+% A similar study is the one done by \cite{bin_ramlan_implementation_2021},
+% where they trained an RL agent to act as an opponent in a fighting game, with the agent being able to adapt to the player's skill level, and provide a more engaging experience, similar to the work done by \cite{vinyals_grandmaster_2019}
 
@@
 
-Independent variables are variables that are manipulated by the researcher, and are mainly used to influence the dependent variables.
-Dependent variables are what happen as a result of the independent variables,
-and are what the researcher is interested in measuring.
+Independent variables are variables that are manipulated by the researcher and are mainly used to influence the dependent variables. 
+Dependent variables are what happen because of the independent variables and are what the researcher is interested in measuring.
 
@@
 
-The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
+The difference between academic and non-academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
 than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
@@
 It is possible to combat this through weakening the player, or making the AI more difficult,
-as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the percieved balance seem unfair, once again breaking the flow \cite{jagdale_finite_2021} \cite{noauthor_finite_2020}.
+as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the perceived balance seem unfair, once again breaking the flow \cite{jagdale_finite_2021} \cite{noauthor_finite_2020}.
 
@@
 due to the complexity of the game, and the many strategies that can be taken, which can be seen in the works done by \cite{vinyals_grandmaster_2019} and \cite{grech_creating_2023}.
-A common tool used in most of these papers is the Unity ML-Agents toolkit, which is a framework for training RL agents in Unity games, making it easier to implement and train RL agents for reasearch purposes.
+A common tool used in most of these papers is the Unity ML-Agents toolkit, which is a framework for training RL agents in Unity games, making it easier to implement and train RL agents for research purposes.
 Along with that, most used the Proximal Policy Optimization (PPO) algorithm, which is a popular RL algorithm that is used for training agents in continuous action spaces, which is especially useful in real-time games.
 
-In the work done by \cite{vinyals_grandmaster_2019}, they trained an RL agent to play StarCraft II, and found that the agent was albe to not only learn how to play the game, but also adapt to the player's actions and strategies in real time.
+In the work done by \cite{vinyals_grandmaster_2019}, they trained an RL agent to play StarCraft II, and found that the agent was able to not only learn how to play the game, but also adapt to the player's actions and strategies in real time.
 A similar result was found in the works done by \cite{bin_ramlan_implementation_2021} and \cite{raut_unity_2024}, where both used the Unity ML-Agents toolkit with PPO to train RL agents in 2 different genres.
@@
 
-\cite{grech_creating_2023} tackled this issue by saving multiple snapshots of the agent during training, and using them to create different difficulties of AI opponents, similarly to how FSM diffifulties are created.
+\cite{grech_creating_2023} tackled this issue by saving multiple snapshots of the agent during training and using them to create different difficulties of AI opponents, similarly to how FSM difficulties are created.
 Their results showed that players reported higher levels of enjoyment and immersion when playing against the RL agents, just like the previous works, with the key difference that new players were also able to enjoy
@@
 to build in a dynamic difficulty adjustment (DDA) system into either the agent or the game, as suggested by \cite{grech_creating_2023}, which would keep the rl agent's difficulty in check, while keeping its
-adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the diffuculty of RL agents.
+adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the difficulty of RL agents.
 
@@
 
+% Takes up too much space, and most are found in the references section anyway.
 \begin{itemize}
+	\item Grech G. (2023). \textit{Creating Difficulty Levels with Reinforcement Learning in a Strategy Game}. \cite{grech_creating_2023}
 	\item Raut, U., Galchhaniya, P., Nehete, A., Shinde, R., \& Bhoite, A. (2024). \textit{Unity ML-Agents: Revolutionizing Gaming Through Reinforcement Learning}. \cite{raut_unity_2024}
@@
 	\item Liu, R.-Z., Pang, Z.-J., Meng, Z.-Y., Wang, W., Yu, Y., \& Lu, T. (2022). \textit{On Efficient Reinforcement Learning for Full-length Game of StarCraft II}. \cite{liu_efficient_2022}
-	\item Abdulghani, A.
+	% \item Abdulghani, A.
 M., Abdulghani, M.
@@
 	\item Vinyals, O., Babuschkin, I., Czarnecki, W. M., et al. (2019). \textit{Grandmaster level in StarCraft II using multi-agent reinforcement learning}. \cite{vinyals_grandmaster_2019}
-	\item Berta, R., Lazzaroni, L., Capello, A., et al. (2024). \textit{Development of Deep-Learning-Based Autonomous Agents for Low-Speed Maneuvering in Unity}. \cite{berta_development_2024}
+	% \item Berta, R., Lazzaroni, L., Capello, A., et al. (2024). \textit{Development of Deep-Learning-Based Autonomous Agents for Low-Speed Maneuvering in Unity}. \cite{berta_development_2024}
 	\item Jagdale, D. (2021). \textit{Finite State Machine in Game Development}. \cite{jagdale_finite_2021}
-	\item \textit{Comparative Analysis of Game Development Techniques: Using Finite State Machine, Physics Simulation, Path Finding, Event Handling}. \cite{noauthor_comparative_nodate}
+	% \item \textit{Comparative Analysis of Game Development Techniques: Using Finite State Machine, Physics Simulation, Path Finding, Event Handling}. \cite{noauthor_comparative_nodate}
 \end{itemize}
@@
 
-The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
-and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
+The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determine their impact on player experience and immersion,
+and what are the key factors that cause this impact/influence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
 and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
@@
 It is important to note that the study
-will also gather a very small amount of qualitative data, making it a \textbf{mixed-methods approach}.
+will also gather an fair amount of qualitative data, making it a \textbf{mixed-methods approach}.
 This is done to help align the more human aspects of player experience with the more quantitative aspects.
@@
 
-% Is more to be done?
-Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
-where the first group will play against the FSM, fill out a questionnaire to gather data on their experience, and then play against the RL agent,
-and again answer the questionnaire. The second group will do the same, but in reverse order, playing against the RL agent first, and then the FSM.
-The end of each experiment's questionnaires will have an open ended question, where players can provide feedback on their experience,
-which will be used to gather qualitative data on their experience. Qualitative data will be used to support the quantitative data gathered from
-the questionnaires, as well as some in-game analytics and logs. This should help capture the more qualitative aspects of player experience,
-while still keeping everything quantitative and measurable. This should help in answering the research questions, and in determining the validity
-of the Hypothesis.
+A simple RTS game was created, using both RL and FSM agents, and a deductive experiment was conducted. The players were split into 2 groups,
+where the first group played against the FSM first, then RL, across all difficulties, and the second group did the same, but with RL first.
+After each difficulty, they were asked to fill out a questionnaire with quantitative and open-ended questions about their experience.
+The qualitative data will help capture the more human aspects of player experience, and along with in-game analytics and logs, 
+will be used to validate the data gathered from the questionnaires. This will help in answering the research questions, and keep everything measurable and comparable.
 
@@
 
-\textbf{Validity}: The study ensures validity by desigining the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
-familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further streghtens the validity of the research being conducted.
+\textbf{Validity}: The study ensures validity by designing the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
+familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further strengthens the validity of the research being conducted.
 
@@
 recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher.
-All datacollection processes will be automated, further standardising the process
+All data collection processes will be automated, further standardising the process
 and ensuring reliability.
@@
 
-\subsection{Etical considerations}
+\subsection{Ethical considerations}
 
@@
 
-% TODO: Discuss the following:
-% You could add a short paragraph explicitly discussing validity (e.g., sample size, playtest design), reliability (consistency of results), and generalizability (applicability to other games or genres).
-% You touch on these in your methodology, but a brief mention here would strengthen the critique.
+The small sample size of participants, along with the results show above indicate that the reliability of the results might be slightly skewed,
+as due to having them replay the game multiple times, they were better able to learn the patterns of the FSM agents, especially on the 3rd playthrough,
+which was the hard difficulty. This, along with the fact that its a single game, could limit the generalisability of the results,
 
@@
 From the analysis of the playtest data, we saw that on easy difficulty, FSM agents provided a more meaningful challenge, and a better player experience than RL agents.
-This could have been due to players not having learned how to exploit the FSM agents yet, or the RL agent being too inconsitent at this difficulty.
+This could have been due to players not having learned how to exploit the FSM agents yet, or the RL agent being too inconsistent at this difficulty.
 As the difficulty increased, RL agents became more competitive, with the average score difference decreasing significantly, and player experience metrics such as engagement and immersion increasing.
@@
 This study set out to compare the impact of Reinforcement Learning (RL) and Finite State Machine (FSM) agents on player experience in Real-Time Strategy (RTS) games.
-By implementing both AI approaches in a custom RTS prototype and conducting a controlled playtest, we gathered quantitative and qualitative data on player performance, engagement,
-and immersion across varying difficulty levels.
+The results from the playtest sessions indicate that in most cases, RL agents provide a more engaging and immersive experience for players.
 
-The results demonstrate that RL agents, particularly at higher difficulties, provide a more engaging and immersive experience for players compared to FSM agents.
-While FSMs can offer a meaningful challenge at lower levels, their predictability becomes apparent as difficulty increases, leading to reduced player engagement.
-In contrast, RL agents adapt to player strategies, maintaining challenge and interest, but can sometimes become too difficult or unpredictable, which may cause frustration for some players.
+The research questions were answered as follows: 
+(1) RL agents generally outperformed FSMs in terms of player engagement at higher difficulty levels, while FSMs were preferred at easier levels; 
+(2) The key factors influencing player experience between the two AI approaches were agent adaptability, predictability, and challenge;
+(3) RL agents greatly improve player immersion, especially as difficulty increases, due to their adaptability and unpredictability, while FSMs suffered from predictability and were exploited by players;
 
-These findings support the hypothesis that RL agents can enhance player experience in RTS games, justifying their higher development and computational costs in contexts where player engagement and
-immersion are priorities. However, careful difficulty balancing is essential to prevent RL agents from overwhelming players, especially less experienced ones.
+As stated in the findings and discussion, the hypothesis was only partially supported, as players reported an improved experience when playing against RL agents
+on all difficulties except for easy, where FSMs were preferred.
 
-% TODO: Discuss the limitations of the study, such as the small sample size, and the fact that it was only conducted on a single game, and not multiple games, or real world games.
+The limited number of participants and a more simplistic game design are the key limitations of this study, as the majority of participants were familiar with the game genre.
+This likely influenced their ability to exploit the FSM agents, and made harder difficulties more desirable, as it better matched their skill level.
+This paired with having to replay the game multiple times on
+harder difficulties may have skewed the results and their reliability.
+As for generalisability, due to it being a small scale study with a single simple game, the results may not be applicable to other RTS games,
+or similar genres.
 
-% TODO: Discuss the implementation of DDA to better close the gap between difficulties, and account for the unpredictability of the RL agent, especially at earlier difficulties where convergence on a strategy
-% is not complete.
-
-% AI suggestions after prompt
-
-% \subsection{Recommendations}
-
-% Based on the findings of this study, the following recommendations are proposed for game developers and researchers:
-
-% \begin{itemize}
-%     \item \textbf{Adopt RL agents for advanced AI in RTS games} where player engagement and adaptability are desired, especially at higher difficulty levels.
-%     \item \textbf{Implement dynamic difficulty adjustment (DDA)} or similar balancing mechanisms to ensure RL agents remain challenging but fair, preventing player frustration.
-%     \item \textbf{Consider hybrid approaches} that combine the predictability and accessibility of FSMs at lower difficulties with the adaptability of RL agents at higher levels,
-% 	catering to a wider range of player skills.
-%     \item \textbf{Further research is needed} to explore long-term player retention, the impact of RL agents in other game genres, and the effectiveness of different balancing strategies.
-%     \item \textbf{Evaluate computational and development costs} in the context of the target audience and desired player experience, as RL agents may not be necessary for all games or player bases.
-% \end{itemize}
-
-% In summary, RL agents have significant potential to improve player experience in RTS games, but their implementation should be carefully managed to maximize benefits and minimize drawbacks.
+For future research, it is recommended to conduct a larger scale study with a more diverse player base, and to integrate Dynamic Difficulty Adjustment (DDA) systems to better balance the difficulty of the game.
+This will help ensure that the game remains challenging across all skill levels, and remove the bias of having players repeat the game multiple times, increasing the difficulty each time.
+Another recommendation is to explore the use of RL agents in other game genres, such as first-person shooters or action-adventure games, to see if the findings of this study can be generalised to other genres.
 
